### 2022-01-24 draft2.tex
add stoke factors for non degenerate case

\DeclareSymbolFont{AMSb}{U}{msb}{m}{n}
\documentclass[11pt,a4paper,twoside,leqno,noamsfonts]{amsart}
           \usepackage{setspace}
\linespread{1.34}           
           %\onehalfspacing
\usepackage[english]{babel}
\usepackage[dvipsnames]{xcolor}
\definecolor{britishracinggreen}{rgb}{0.0, 0.26, 0.15}
\definecolor{cobalt}{rgb}{0.0, 0.28, 0.67}
\usepackage[utopia]{mathdesign}
    \DeclareSymbolFont{usualmathcal}{OMS}{cmsy}{m}{n}
    \DeclareSymbolFontAlphabet{\mathcal}{usualmathcal}
\usepackage[a4paper,top=4cm,bottom=3cm,left=3.5cm,
           right=3.5cm,bindingoffset=5mm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{braket,caption,comment,mathtools,stmaryrd}
\usepackage{multirow,booktabs,microtype}
\usepackage{latexsym}
\usepackage{todonotes}
\usepackage{fancyhdr}
%\renewcommand{\sectionmark}[1]{\markboth{\thesection\ #1}{}}
\pagestyle{fancy}
% Clear the header and footer
\fancyhead{}
\fancyfoot{}
% Set the right side of the footer to be the page number
\fancyfoot[R]{\thepage}
\addtolength{\headheight}{\baselineskip}
%\fancyhead[RE]{\rightmark}
%\fancyhead[RE]{}
\usepackage{soul} % per testo barrato
\usepackage[colorlinks,bookmarks]{hyperref} %
\hypersetup{colorlinks,%
            citecolor=britishracinggreen,%
            filecolor=black,%
            linkcolor=cobalt,%
            urlcolor=black}
\setcounter{tocdepth}{2}
%\setcounter{section}{-1}
\numberwithin{equation}{section}
\input{command}
\DeclareRobustCommand{\subtitle}[1]{\\#1}
\title[Exponential Integrals]{Exponential Integrals\\ [1ex]
  }

\author{
Veronica Fantini 
}
%\address{SISSA Trieste, Via Bonomea 265, 34136 Trieste, Italy}
%\email{ vfantini@sissa.it}
%
\begin{document}
\hbadness=150
\vbadness=150
%
%\begin{abstract}
%We show how wall-crossing formulas in coupled $2d$-$4d$ systems, introduced by Gaiotto, Moore and Neitzke, can be interpreted geometrically in terms of the deformation theory of holomorphic pairs, given by a complex manifold together with a holomorphic vector bundle. The main part of the paper studies the relation between scattering diagrams and deformations of holomorphic pairs, building on recent work by Chan, Conan Leung and Ma.      
%\end{abstract}
%
\maketitle
%
%
%
{
\hypersetup{linkcolor=black}
\tableofcontents}

\section{Introduction}

%\begin{example}[Airy]
%The Airy equation is 
%\begin{equation}
%y''=xy
%\end{equation}
% and a system of solutions is given by (see DLMF) 
%
%\begin{align*}
%&Ai(x)=\frac{1}{2\pi i}\int_{\infty e^{-i\pi/3}}^{\infty e^{i\pi/3}}e^{t^3/3-xt}dt\\
%&Bi(x)=\frac{1}{2\pi}\int_{-\infty}^{\infty e^{-i\pi/3}}e^{t^3/3-xt}dt+\frac{1}{2\pi}\int_{-\infty}^{\infty e^{i\pi/3}}e^{t^3/3-xt}dt
%\end{align*}
%
%
%
%Let $\psi(x)\defeq\sum_{n\geq 0}a_nx^{-n}$ be the asymptotic expansion of $Ai(x)$ as $x\to\infty$: from DLMF and Sauzin
%
%\begin{equation}
%Ai(x)\sim \frac{e^{-2/3x^{3/2}}}{2\sqrt{\pi}x^{1/4}}\sum_{n\geq 0}(-1)^n\frac{1}{2^{n+1}\pi n!}\Gamma\left(n+\frac{5}{6}\right)\Gamma\left(n+\frac{1}{6}\right)\left(\frac{3}{2}x^{-3/2}\right)^n
%\end{equation}
%and $a_n\defeq (-1)^n\frac{1}{2\pi n!}\Gamma\left(n+\frac{5}{6}\right)\Gamma\left(n+\frac{1}{6}\right)\left(\frac{3}{4}\right)^n$. Analougusly, let $\varphi(x)\defeq\sum_{n\geq 0}b_nx^{-n}$ be the asymptotic expansion of $Bi(x)$ as $x\to\infty$: from DLMF and Sauzin
%
%\begin{equation}
%Bi(x)\sim \frac{e^{2/3x^{3/2}}}{\sqrt{\pi}x^{1/4}}\sum_{n\geq 0}\frac{1}{2^{n+1}\pi n!}\Gamma\left(n+\frac{5}{6}\right)\Gamma\left(n+\frac{1}{6}\right)\left(\frac{3}{2}x^{-3/2}\right)^n
%\end{equation}
%
%and $b_n\defeq \frac{1}{2\pi n!}\Gamma\left(n+\frac{5}{6}\right)\Gamma\left(n+\frac{1}{6}\right)\left(\frac{3}{4}\right)^n$.
%\begin{prop}
%$\psi(x)$ and $\varphi(x)$ are resurgent series.
%\end{prop}
%\begin{proof}
%Let $\hat{\psi}(\zeta)=\frac{1}{2\pi}\sum_{n\geq 0}a_{n-1}\frac{\zeta^n}{(n-1)!}$, mathematica says 
%\begin{equation}
%\hat{\psi}(\zeta)=-\frac{5}{48} F_1(7/6,11/6,2,-3/4\zeta)
%\end{equation}
%which has a log singularity at $\zeta=-4/3$, namely 
%\begin{equation}
%\hat{\psi}(\zeta+4/3)=\log(\zeta)\hat{\psi}_{-4/3}(\zeta)+\text{hol.fct}
%\end{equation}
%where $\hat{\psi}_{-4/3}(\zeta)$ is holomorphic in a neighbourhood of the origin. In particular, $\hat{\psi}_{-4/3}(\zeta)=-\frac{1}{2}\varphi(\zeta)$.    
%Now, let $\hat{\varphi}(\zeta)=\frac{1}{2\pi}\sum_{n\geq 0}b_{n-1}\frac{\zeta^n}{(n-1)!}$, mathematica says 
%\begin{equation}
%\hat{\varphi}(\zeta)=\frac{5}{48} F_1(7/6,11/6,2,3/4\zeta)
%\end{equation}
%which has a log singularity at $\zeta=4/3$, namely 
%\begin{equation}
%\hat{\varphi}(\zeta-4/3)=\log(\zeta)\hat{\varphi}_{4/3}(\zeta)+\text{hol.fct}
%\end{equation}
%where $\hat{\varphi}_{4/3}(\zeta)$ is holomorphic in a neighbourhood of the origin. In particular, $\hat{\varphi}_{4/3}(\zeta)=\frac{1}{2}\psi(\zeta)$.    
%\end{proof}
%
%Comment on the asymptotic expansion of $Ai(x)$ and $Bi(x)$: the prepared form of the Airy equation is obtained after the change of coordinates $z=x^{3/2}$ and it reads
%
%\begin{equation}\label{prepform}
%y''(z)+\frac{y'(z)}{3z}=\frac{4}{9}y(z)
%\end{equation} 
%
%In particular, its Borel trasform is 
%\begin{equation}
%\zeta^2\hat{y}-\frac{1}{3}\ast (\zeta\hat{y})=\frac{4}{9}\hat{y}
%\end{equation}
%and we read the singularity of the Borel transform as solution of $z^2-4/9=0$, which are $z=\pm 2/3$. These are indeed the exponent of the exponential factor in the asymptotic of $Ai(z)$ and $Bi(z)$ respectively.  
%
%Alternatively, from the theory of ODE we can look for a solution of \eqref{prepform} of the form
%\begin{equation}
%y(z)=\sum_{k\in\Z^2}U^ke^{-k\cdot\lambda z}z^{-\tau\cdot k}w_{k}(z)
%\end{equation} 
%where $\lambda=(\lambda_1,\lambda_2)$ with $\lambda_i^2-4/9=0$, $\tau=(\tau_1,\tau_2)$ with $\tau_i=1/6$ and $w_k(z)\in\C[[z^{-1}]]$ is a formal series. Plugging in the ansatz in \eqref{prepform}, we fins that the only non zero solutions $w_k$ occurs for $k=(1,0)$ and $k=(0,1)$, meaning that the formal integral of the Airy equation is given by 
%\begin{equation}
%y(z)=U_1e^{-2/3z}z^{-1/6}w_{(1,0)}(z)+U_2e^{2/3z}z^{-1/6}w_{(0,1)}(z).
%\end{equation}
%The latter expression can be compared with the asymptotic behaviour of $Ai(z)+Bi(z)$ as $z\to\infty$. In particular, $w_{k}$ solves the following homogeneus equation:
%\begin{align*}
%\left(\partial_z^2-\frac{4}{3}\partial_z+\frac{5}{36z^2}\right)w_{(1,0)}=0\\
%\left(\partial_z^2+\frac{4}{3}\partial_z+\frac{5}{36z^2}\right)w_{(0,1)}=0
%\end{align*}
%Assuming $w_{(1,0)}=1+\sum_{n\geq 1}a_n^+z^{-n}$ and $w_{(0,1)}=1+\sum_{n\geq 1}a_n^-z^{-n}$ we get \[a_n^+=(-1)^n\left(\frac{3}{4}\right)^n\frac{\Gamma(n+\frac{1}{6})\Gamma(n+\frac{5}{6})}{2\pi n!}=a_n\] and \[a_n^-=\left(\frac{3}{4}\right)^n\frac{\Gamma(n+\frac{1}{6})\Gamma(n+\frac{5}{6})}{2\pi n!}=b_n.\] 
%
%Now, we investigate the exponential integral
%\begin{equation}
%I(z)\defeq\int_{\Gamma} \exp(-z\left(\frac{x^3}{3}-x\right))dx
%\end{equation}
%where $\Gamma$ is a suitable countour. By simple computation we notice that $I(z)=-2\pi iz^{-1/3}Ai(x)$ (where $z=x^{3/2}$). In particular, $I(z)$ is a solution of 
%\begin{equation}\label{eq:Iold}
%I''-\frac{4}{9}I+\frac{I'}{z}-\frac{1}{9}\frac{I}{z^2}=0
%\end{equation} 
% The Borel trasform of $\hat{I}(z)$ can be computed directly from the equation \eqref{eq:I}, namely the Borel transform of equation \eqref{eq:I} is
%\begin{equation}
%\zeta^2 \hat{I}-\frac{4}{9}\hat{I}-1\ast\zeta\hat{I}-\frac{1}{9}\zeta\ast\hat{I}=0
%\end{equation} 
%\begin{equation}
%\zeta^2 \hat{I}-\frac{4}{9}\hat{I}-\int_0^\zeta\zeta'\hat{I}(\zeta')d\zeta '-\frac{1}{9}\int_0^\zeta(\zeta-\zeta')\hat{I}(\zeta')d\zeta'=0
%\end{equation} 
%\begin{align*}
%2\zeta\hat{I}+\zeta^2\hat{I}'-\frac{4}{9}\hat{I}'-\zeta\hat{I}-\frac{1}{9}\int_0^\zeta\hat{I}(\zeta')d\zeta'=0\\
%\hat{I}+\zeta\hat{I}'+2\zeta\hat{I}'+\zeta^2\hat{I}''-\frac{4}{9}\hat{I}''-\frac{1}{9}\hat{I}=0
%\end{align*}
%\begin{equation}
%(\zeta^2-\frac{4}{9})\hat{I}''+3\zeta\hat{I}'+(1-\frac{1}{9})\hat{I}=0
%\end{equation} 
%and for some costants $c_1,c_2\in\C$ the solutions are
%\begin{equation}
%\hat{I}(\zeta)=\frac{9\zeta c_1}{4\sqrt{\frac{4}{9}-\zeta^2}}+c_2
%\end{equation}  
%\end{example}
%\textcolor{red}{Is $F_1(7/6,11/6,2,3/4\zeta)$ algebraic?}

\section{Fractional derivatives and Borel transform}

\begin{definition}
Let $\alpha\in (0,1)$ and $n\in\N$, then the $n+\alpha$-Caputo's derivative of a smooth fucntion $f$ is defined as
\begin{equation}
\partial_x^{n+\alpha}f(x)\defeq\frac{1}{\Gamma(1-\alpha)}\int_0^x(x-s)^{-\alpha}f^{(n+1)}(s)ds
\end{equation}
\end{definition}

In particular, this definition is well suited for the differential calculus in the convolutive model $\left(\C[\![\zeta]\!], \ast\right)$. Let $\varphi(z)\defeq\sum_{k\geq 0}a_kz^{-k-1}\in\C[\![z^{-1}]\!]$ be Gevrey $1$, then assuming $a_k=0$ for every $k<n$, the Borel transform of $z^{n+\alpha}\varphi(z)$ can be computed in two different ways:
\begin{multline}
\label{mod1}
\mathcal{B}\left(z^{n+\alpha}\varphi(z)\right)(\zeta)=\mathcal{B}(z^{\alpha+n})\ast\hat{\varphi}(\zeta)=\int_0^{\zeta}\frac{(\zeta-s)^{-1-n-\alpha}}{(-1-n-\alpha)!}\sum_{k\geq 0}\frac{a_k}{k!}s^{k}ds\\
=\frac{1}{(-\alpha)!}\int_0^{\zeta}(\zeta-s)^{-\alpha}\sum_{k\geq 0}\frac{a_k}{(k-n-1)!}s^{k-n-1}ds=\partial_{\zeta}^{n+\alpha}\hat{\varphi}(\zeta)
\end{multline}
\begin{equation}
\label{mod2}
\mathcal{B}\left(z^{n+\alpha}\varphi(z)\right)(\zeta)=\mathcal{B}\left(\sum_{k\geq 0}a_kz^{-k-1+n+\alpha}\right)(\zeta)=\sum_{k>n}\frac{a_k}{(k-n-\alpha)!}\zeta^{k-n-\alpha}
\end{equation}

and computitng the integral which defines the $n+\alpha$-derivative in \eqref{mod1} we get exaclty the same result as \eqref{mod2}.

\section{Resurgence of exponential integrals}

Let $X$ be a $N-\dim$ manifold, $f\colon X\to\C$ be a holomorphic Morse function with only simple critical points, and $\nu\in\Gamma(X,\Omega^N)$, and set

\begin{equation}
I(z)\defeq\int_{\mathcal{C}}e^{-zf}\nu
\end{equation}
where $\mathcal{C}$ is a suitable countur such that the integral is well defined.  
For any Morse cirtial points $x_\alpha$ of $f$, the saddle point approximation gives the following formal series 
\begin{equation}
I_{\alpha}(z)\defeq\int_{\mathcal{C}_\alpha}e^{-zf}\nu\sim \tilde{I}_{\alpha}\defeq e^{-zf(x_\alpha)}(2\pi)^{N/2} z^{-N/2}\sum_{n\geq 0}a_{\alpha,n}z^{-n} \qquad \text{ as } z\to\infty
\end{equation}
where $\mathcal{C}_\alpha$ is a steepest descendet path through the critical point $x_\alpha$. 
\begin{theorem}\label{thm:maxim} Let $N=1$. Let $\tilde{\varphi}_{\alpha}(z)\defeq e^{-zf(x_\alpha)}(2\pi)^{1/2} \sum_{n\geq 0}a_{\alpha,n}z^{-n}$ and assume $f''(x_\alpha)\neq 0$ for every critical point $x_\alpha$. Then
\begin{enumerate}
\item $\tilde{\varphi}_\alpha$ is Gevrey-1;
\item $\hat{\varphi}(\zeta)\defeq\mathcal{B}(\tilde{\varphi})$ is a germ of analytic function at $\zeta=\zeta_{\alpha}=f(x_\alpha)$;
\item the following formual holds true
\begin{multline}\label{formula1}
\hat{\varphi}_{\alpha}(\zeta)=\partial^{\textcolor{red}{3/2}}_{\zeta, \text{based at }\zeta_\alpha}\left(\int_{f^{-1}(\zeta_\alpha)}^{f^{-1}(\zeta)}\nu\right)=\Gamma(-1/2)\int_{\zeta_\alpha}^\zeta (\zeta-\zeta')^{-1/2}\textcolor{red}{\partial_{\zeta'}\left(\int_{f^{-1}(\zeta')}\frac{\nu}{df}\right)} d\zeta'
\end{multline} 
\end{enumerate}
\end{theorem}
\begin{proof}
Part $(1)$: Let us assume that locally $\nu=\sum_{j\geq 0}b_j^{\alpha}(t-t_{\alpha})^{j}dt$ for $|t-t_\alpha|<\varepsilon$. By the steepest descendet method, $I_\alpha(z)$ can be aproximated as $z\to\infty$ as
\begin{align*}
I_\alpha(\zeta)&\sim e^{-zf(x_\alpha)}\int_{-\varepsilon}^{\varepsilon}\sum_{n\geq 0}t^{2n}b_{2n}^{\alpha}e^{-zf''(x_\alpha)\frac{t^2}{2}}dt\\
&=e^{-zf(x_\alpha)}\sum_{n\geq 0}b_{2n}^{\alpha}\int_{-\varepsilon}^{\varepsilon}t^{2n}e^{-zf''(x_{\alpha})\frac{t^2}{2}}dt\\
&=\sqrt{2\pi}e^{-zf(x_\alpha)}z^{-1/2}\sum_{n\geq 0}\frac{b_{2n}^{\alpha}}{(f''(x_{\alpha}))^{n+1/2}}(2n-1)!! z^{-n}\mathrm{Erf}\left(\frac{\sqrt{z f''(x_\alpha)}}{\sqrt{2}}\varepsilon\right)+\\
&-2e^{-zf(x_\alpha)}\sum_{n\geq 1}b_{2n}^{\alpha}(2n-1)!!e^{-zf''(x_{\alpha})\frac{\varepsilon^2}{2}}\sum_{j=1}^n\frac{\varepsilon^{2j-1}}{(2j-1)!!}\left(f''(x_\alpha)z\right)^{n-j+1}\\
&\sim_{\varepsilon<<1} 2\varepsilon \sqrt{2\pi}e^{-zf(x_\alpha)}z^{-1/2}\sum_{n\geq 0}\frac{b_{2n}^{\alpha}}{(f''(x_{\alpha}))^{n+1/2}}(2n-1)!! z^{-n}
\end{align*} 
therefore \[a_{\alpha,n}\defeq 2\varepsilon\frac{b_{2n}^{\alpha}}{(f''(x_{\alpha}))^{n+1/2}}(2n-1)!!\]

Since, $\tfrac{1}{\varepsilon}=\limsup_{n}\sqrt[n]{b_{n}^{\alpha}}$

\[|a_{\alpha,n}|\leq CA^n n!\]
where we use $(2n-1)!!\sim \tfrac{2^n}{\sqrt{\pi n}}n!$ as $n\to \infty$.

Part $(2)$: \begin{align*}
\hat{\varphi}_\alpha(\zeta)=\mathcal{B}\left(e^{-zf(x_\alpha)}(2\pi)^{1/2} \sum_{n\geq 0}a_{\alpha,n}z^{-n}\right)(\zeta)=T_{f(x_\alpha)}(2\pi)^{1/2} \left(\delta a_0+\sum_{n\geq 0}a_{n+1}\frac{\zeta^n}{n!}\right)\\
(2\pi)^{1/2} \left(\delta(f_{x_\alpha}) a_0+\sum_{n\geq 0}a_{n+1}\frac{(\zeta-f(x_\alpha))^n}{n!}\right)
\end{align*}
Since $a_{n}\leq CA^nn!$, the series $\sum_{n\geq 0}a_{n+1}\frac{(\zeta-f(x_\alpha))^n}{n!}$ has a finite radius of convergence.

Part $(3)$: thanks to properties of Caputo's fractional derivatives, we have that the Borel transform of $\tilde{I}_{\alpha}(z)=z^{-1/2}\tilde{\varphi}_\alpha(z)$ is 
\begin{equation}
\partial^{1/2}_{\zeta, \text{based at }\zeta_\alpha}\hat{I}_{\alpha}(\zeta)=\hat{\varphi}_{\alpha}(\zeta).
\end{equation} 
In addition, we notice 
\begin{align*}
I_{\alpha}(z)&=\int_{\mathcal{C}_{\alpha}}e^{-zf}\nu & f=\zeta\\
&=\int_{\mathcal{H}_{\alpha}}e^{-z\zeta}\partial_{\zeta}\left(\int_{f^{-1}(\zeta_\alpha)}^{f^{-1}(\zeta)}\nu\right)d\zeta   & \mathcal{H}_{\alpha} \text{ is Hankel countour} \\
&=:\int_{\mathcal{H}_{\alpha}}e^{-\zeta z}\hat{I}_{\alpha}(\zeta)d\zeta &
\end{align*}
hence 
\begin{equation}
\hat{\varphi}_{\alpha}(\zeta)=\partial^{1/2}_{\zeta, \text{based at } \zeta_\alpha}\left(\partial_{\zeta}\left(\int_{f^{-1}(\zeta_\alpha)}^{f^{-1}(\zeta)}\nu\right)\right)=\partial^{3/2}_{\zeta, \text{based at } \zeta_\alpha}\left(\int_{f^{-1}(\zeta_\alpha)}^{f^{-1}(\zeta)}\nu\right)
\end{equation}
\end{proof}

\begin{example}[Airy]
Let $f(t)=\frac{t^3}{3}-t$ and \[I(z)\defeq\int_{\gamma} e^{-zf(t)}\text{dt}\]
where $\gamma$ is a countour where the integral is well defined.

By the change of coordinates $z=x^{3/2}$, $I(z)=-2\pi iz^{-1/3}Ai(x)$ where \[Ai(x)\defeq \frac{1}{2\pi i}\int_{-\infty e^{-i\frac{\pi}{3}}}^{\infty e^{i\frac{\pi}{3}}} e^{\frac{t^3}{3}-zt}\text{dt}\] 
hence $I(z)$ solves the following ODE\footnote{$Ai(x)$ solves the Airy equation $y''=xy$.}
\begin{equation}\label{eq:I}
I''(z)-\frac{4}{9}I(z)+\frac{I'(z)}{z}-\frac{1}{9}\frac{I(z)}{z^2}=0
\end{equation}
A formal solution of \eqref{eq:I} can be computed by making the following ansatz 
\begin{equation}
\tilde{I}(z)=\sum_{k\in\N^2}U^ke^{-\lambda\cdot k z}z^{-\tau\cdot k}w_k(z)
\end{equation}
with $U^{(k_1,k_2)}=U_1^{k_1}U_2^{k_2}$ and $U_1,U_2\in\C$ are constant parameter, $\lambda=(\frac{2}{3},-\frac{2}{3})$, $\tau=(\frac{1}{2},\frac{1}{2})$, and $\tilde{w}_k(z)\in\C[[z^{-1}]]$. In addition, we can check that the only non zero $\tilde{w}_k(z)$ occurs at $k=(1,0)$ and $k=(0,1)$, therefore
\begin{equation}
\tilde{I}(z)=U_1e^{-2/3z}z^{-1/2}\tilde{w}_{+}(z)+U_2e^{2/3z}z^{-1/2}\tilde{w}_{-}(z)
\end{equation}  
where from now on we denote $\tilde{w}_+(z)=\tilde{w}_{(1,0)}(z)$ and $\tilde{w}_-(z)=\tilde{w}_{(0,1)}(z)$. In particular, $\tilde{w}_+(z)$ and $\tilde{w}_-(z)$ are formal solution of 
\begin{align}
\label{eq:w+} \tilde{w}_+''-\frac{4}{3}\tilde{w}_+'+\frac{5}{36}\frac{\tilde{w}_+}{z^2}=0\\
\label{eq:w-} \tilde{w}_-''+\frac{4}{3}\tilde{w}_-'+\frac{5}{36}\frac{\tilde{w}_-}{z^2}=0
\end{align}
Taking the Borel transform of \eqref{eq:w+}, \eqref{eq:w-} we get
\begin{align*}
&\zeta^2\hat{w}_{+}(\zeta)+\frac{4}{3}\zeta\hat{w}_{+}+\frac{5}{36}\zeta\ast\hat{w}_{+}=0\\
&\zeta^2\hat{w}_{+}(\zeta)+\frac{4}{3}\zeta\hat{w}_{+}+\frac{5}{36}\int_0^\zeta(\zeta-\zeta')\hat{w}_{+}(\zeta')d\zeta'=0
\end{align*}
\begin{align*}
&\zeta^2\hat{w}_{-}(\zeta)-\frac{4}{3}\zeta\hat{w}_{-}+\frac{5}{36}\zeta\ast\hat{w}_{-}=0\\
&\zeta^2\hat{w}_{-}(\zeta)-\frac{4}{3}\zeta\hat{w}_{-}+\frac{5}{36}\int_0^\zeta(\zeta-\zeta')\hat{w}_{-}(\zeta')d\zeta'=0
\end{align*}
and taking derivatives we get
\begin{align*}
&\zeta(\frac{4}{3}+ \zeta)\hat{w}_{+}''+(\frac{8}{3}+4\zeta)\hat{w}_+'+\frac{77}{36}\hat{w}_{+}=0 & \\
&\frac{4}{3}\zeta( 1+ \frac{3}{4}\zeta)\hat{w}_{+}''+(\frac{8}{3}+4\zeta)\hat{w}_+'+\frac{77}{36}\hat{w}_{+}=0 & \\
&\qquad u(1-u)\hat{w}_+''(u)+(2-4u)\hat{w}_+'(u)-\frac{77}{36}\hat{w}_+(u)=0 & u=-\frac{3}{4}\zeta\\
%&\qquad u(1-u)\hat{w}_-''(u)+(2-4u)\hat{w}_-'(u)-\frac{77}{36}\hat{w}_-(u) & u=\frac{3}{4}\zeta
\end{align*} 
\begin{align*}
&\zeta(-\frac{4}{3}+ \zeta)\hat{w}_{-}''+(-\frac{8}{3}+4\zeta)\hat{w}_-'+\frac{77}{36}\hat{w}_{-}=0 & \\
&\frac{4}{3}\zeta(-1+ \frac{3}{4}\zeta)\hat{w}_{-}''+(-\frac{8}{3}+4\zeta)\hat{w}_-'+\frac{77}{36}\hat{w}_{-}=0 & \\
&\qquad u(1-u)\hat{w}_-''(u)+(2-4u)\hat{w}_-'(u)-\frac{77}{36}\hat{w}(u)=0 & u=\frac{3}{4}\zeta
\end{align*} 
Notice that the latter equations are hypergeometric, hence a solution is given by 
\begin{align}
\label{hat+}\hat{w}_+(\zeta)=\mathit{c}_1 \, {}_{1}F_{2}\left(\frac{7}{6},\frac{11}{6},2,-\frac{3}{4}\zeta\right)\\
\label{hat-}\hat{w}_-(\zeta)=\mathit{c}_2 \, {}_{1}F_{2}\left(\frac{7}{6},\frac{11}{6},2,\frac{3}{4}\zeta\right)
\end{align}
for some constants $\mathit{c}_1, \mathit{c}_2\in\C$ (see DLMF 15.10.2). In addition $\hat{w}_{\pm}(\zeta)$ have a log singularity respectively at $\zeta=\mp\frac{4}{3}$, therefore they are $\lbrace\mp\frac{4}{3}\rbrace $-resurgent functions.\footnote{The solution we find are equal to the ones in DLMF $\mathsection 9.7$. They also agree with slide 10 of Maxim's talk for ERC and with the series (6.121) in Sauzin's book. However they do not agree with (2.16) in Mari\~no's Diablerets.}

\begin{remark}
$\hat{w}_{+}(\zeta)$ is Laplace summable along the positive real axis, and it can be analyticaly continued on $\C\setminus -\frac{4}{3}\R_{\leq 0}$ with (see 15.2.3 DLMF)
\begin{align*}
\hat{w}_+(\zeta+i0)-\hat{w}_+(\zeta-i0)&=-\frac{36}{5}i(-\frac{3}{4}\zeta-1)^{-1}\sum_{n\geq 0}\frac{(5/6)_n(1/6)_n}{\Gamma(n)n!}(1+\frac{3}{4}\zeta)^n &\zeta<-\frac{4}{3}\\
&=\frac{36}{5}i\sum_{n\geq 0}\frac{(5/6)_n(1/6)_n}{\Gamma(n)n!}(1+\frac{3}{4}\zeta)^{n-1} & \\
&=-\frac{36}{5}i(-\frac{3}{4}\zeta-1)^{-1}\left(\frac{5}{144} (4 + 3 \zeta)\left(1+{}_{1}F_{2}\left(\frac{7}{6},\frac{11}{6},2,1+\frac{3}{4}\zeta\right)\right)\right) &\\
&=\mathbf{i}\,\,{}_{1}F_{2}\left(\frac{7}{6},\frac{11}{6},2,1+\frac{3}{4}\zeta\right) &\\
&=\mathbf{i}\hat{w}_{-}(\zeta+\frac{4}{3}) &
\end{align*}
Anolougusly, $\hat{w}_-(\zeta)$ is Laplace summable along the negative real axis, and it jumps across the branch cut $\frac{4}{3}\R_{\geq 0}$ as 
\begin{align*}
\hat{w}_-(\zeta+i0)-\hat{w}_-(\zeta-i0)&=\frac{36}{5}i(\frac{3}{4}\zeta-1)^{-1}\sum_{n\geq 0}\frac{(5/6)_n(1/6)_n}{\Gamma(n)n!}(1-\frac{3}{4}\zeta)^n & \zeta>\frac{4}{3}\\
&=\frac{36}{5}i(\frac{3}{4}\zeta-1)^{-1}\left(-\frac{5}{144}(-4+3\zeta){}_{1}F_{2}\left(\frac{7}{6},\frac{11}{6},2,1-\frac{3}{4}\zeta\right)\right)  &\\
&=-\mathbf{i}\pi\,\,{}_{1}F_{2}\left(\frac{7}{6},\frac{11}{6},2,1-\frac{3}{4}\zeta\right) &\\
&=-\mathbf{i}\hat{w}_{+}(\zeta-\frac{4}{3}) &
\end{align*}
These relations manifest the resurgence property of $\tilde{I}$, indeed near the singularities in the Borel plane of either $\hat{w}_+$ or $\hat{w}_-$, $\hat{w}_-$ and $\hat{w}_+$ respectively contribute to the jump of the former solution.  
\end{remark}

%\begin{remark}
%In \cite{Marino-diableret}, Mari\~{n}o studies the example of the Airy function and its resurgent properties. According to his notation
%\begin{equation}
%Ai(x)=\frac{1}{2x^{1/4}\sqrt{\pi}}e^{-\frac{2}{3}x^{3/2}}\varphi_1(x^{-3/2})
%\end{equation}
%\begin{equation}
%Bi(x)=\frac{1}{2x^{1/4}\sqrt{\pi}}e^{\frac{2}{3}x^{3/2}}\varphi_2(x^{-3/2})
%\end{equation}
%with \[\varphi_{1,2}(z)=\sum_{k\geq 0}\frac{1}{2\pi}\left(\mp\frac{3}{4}\right)^k\frac{\Gamma\left(k+\frac{1}{6}\right)\Gamma\left(k+\frac{5}{6}\right)}{k!}z^k.\]
%In particular, the Borel transform of $\varphi_1(z)$ and $\varphi_2(z)$ are respectively 
%\begin{align}
%\label{Marino_hat}
%\hat{\varphi}_1(\zeta)&={}_1F_2\left(\frac{1}{6},\frac{5}{6};1;-\frac{3}{4}\zeta\right)\\
%\label{Marino_hat2}\hat{\varphi}_2(\zeta)&={}_1F_2\left(\frac{1}{6},\frac{5}{6};1;\frac{3}{4}\zeta\right)
%\end{align} 
% In particular, we notice that $\hat{w}_+(\zeta)$ and $\hat{w}_-(\zeta)$ in \eqref{eq:w+}, \eqref{eq:w-} are (up to a constant) the first derivatives of
%\begin{align*}
%\hat{w}_+(\zeta)&\propto\frac{\dd}{\dd\zeta}\hat{\varphi}_1(\zeta)\\
%\hat{w}_-(\zeta)&\propto\frac{\dd}{\dd\zeta}\hat{\varphi}_2(\zeta)
%\end{align*} 
%\end{remark}

Our next goal is to prove that the Borel transform of $\tilde{I}(z)$ can be written in terms of $1/f'(f^{-1}(\zeta))$, namely formula \eqref{formula1}. It is convenient to consider the two asymptotic formal solutions separately, namely we define

\begin{align}
\tilde{I}_{-1}(z)\defeq e^{-2/3z}z^{-1/2}\tilde{w}_+(z)=\colon z^{-1/2}\tilde{u}_+(z) \\
\tilde{I}_{1}(z)\defeq e^{2/3z}z^{-1/2}\tilde{w}_-(z)=\colon z^{-1/2}\tilde{u}_-(z)
\end{align}

In particular, $\tilde{u}_{\pm}(z)$ are solutions of

\begin{equation}\label{eq:u}
\tilde{u}''(z)-\frac{4}{9}\tilde{u}(z)+\frac{5}{36}\frac{\tilde{u}(z)}{z^2}=0
\end{equation}

with asymptotic behaviour $\tilde{u}_\pm(z)\sim O(e^{\pm 2/3 z})$ as $z\to\infty$.

The Borel transforms $\hat{u}_{\pm}(\zeta)$ solve the same equation
\begin{align*}
&\zeta^2\hat{u}-\frac{4}{9}\hat{u}+\frac{5}{36}\zeta\ast\hat{u}\\
&\zeta^2\hat{u}-\frac{4}{9}\hat{u}+\frac{5}{36}\int_0^\zeta(\zeta-\zeta')\hat{u}(\zeta')d\zeta'\\
&\text{taking derivatives is equivalent to} \\
&(\zeta^2-\frac{4}{9})\hat{u}''(\zeta)+4\zeta\hat{u}'(\zeta)+\frac{77}{36}\hat{u}(\zeta)=0
\end{align*}

and Mathematica gives the following solutions
\begin{align*}
\hat{u}(\zeta)&=c_1 \, {}_{1}F_{2}\left(\frac{7}{12},\frac{11}{12},\frac{1}{2},\frac{9}{4}\zeta^2\right) +\frac{3i}{2}\zeta c_2 \, {}_{1}F_{2}\left(\frac{13}{12},\frac{17}{12},\frac{3}{2},\frac{9}{4}\zeta^2\right)= &\\
&=c_1\frac{\Gamma(\frac{13}{12})\Gamma(\frac{17}{12})}{2\sqrt{\pi}}\left({}_{1}F_{2}\left(\frac{7}{12},\frac{11}{12},\frac{1}{2},\frac{1}{2}-\frac{3}{4}\zeta\right)- {}_{1}F_{2}\left(\frac{7}{12},\frac{11}{12},\frac{1}{2},\frac{1}{2}+\frac{3}{4}\zeta\right)\right) & \text{see DLMF 15.8.27} \\
&\qquad +\frac{3i}{2}\zeta c_2 \left(\frac{\Gamma(\frac{7}{12})\Gamma(\frac{11}{12})}{3\zeta\Gamma(-\frac{1}{2})\Gamma(2)}\right)\left( {}_{1}F_{2}\left(\frac{7}{6},\frac{11}{6},{2},\frac{1}{2}-\frac{3}{4}\zeta\right)-{}_{1}F_{2}\left(\frac{7}{6},\frac{11}{6},{2},\frac{1}{2}+\frac{3}{4}\zeta\right)\right)  & \text{see DLMF 15.8.28}\\
&=\left(c_1\frac{\Gamma(\frac{13}{12})\Gamma(\frac{17}{12})}{2\sqrt{\pi}} -c_2i\frac{\Gamma(\frac{7}{12})\Gamma(\frac{11}{12})}{4\sqrt{\pi}}\right)\, {}_{1}F_{2}\left(\frac{7}{6},\frac{11}{6},2,\frac{1}{2}-\frac{3}{4}\zeta\right)  + & \\
&\qquad\qquad +\left(c_1\frac{\Gamma(\frac{13}{12})\Gamma(\frac{17}{12})}{2\sqrt{\pi}} +c_2i\frac{\Gamma(\frac{7}{12})\Gamma(\frac{11}{12})}{4\sqrt{\pi}}\right)\, {}_{1}F_{2}\left(\frac{7}{6},\frac{11}{6},2,\frac{1}{2}+\frac{3}{4}\zeta\right) &  
\end{align*}
Since $\hat{u}_+$ has a simple singularity at $\zeta=-2/3$ and $\hat{u}_-$ has a simple singularity at $\zeta=2/3$, we have
\begin{align*}
\hat{u}_+(\zeta)&=C_1 T_{-2/3} {}_{1}F_{2}\left(\frac{7}{6},\frac{11}{6},2,-\frac{3}{4}\zeta\right)=C_1 T_{-2/3}\hat{w}_+(\zeta)\\
\hat{u}_-(\zeta)&= C_2 T_{2/3} {}_{1}F_{2}\left(\frac{7}{6},\frac{11}{6},2,\frac{3}{4}\zeta\right)= C_2 T_{2/3}\hat{w}_-(\zeta) & \\
%&=C_3 T_{-2/3}\textcolor{red}{\left(\frac{3}{2\sqrt{\zeta}}\right)^{-1}\partial^{1/2}_{\zeta} {}_{1}F_{2}\left(\frac{2}{3},\frac{4}{3},\frac{3}{2},-\frac{3}{4}\zeta\right)}+ C_4\textcolor{red}{\left(\frac{3}{2\sqrt{\zeta}}\right)^{-1}\partial^{1/2}_\zeta {}_{1}F_{2}\left(\frac{2}{3},\frac{4}{3},\frac{3}{2},\frac{3}{4}\zeta\right)}
\end{align*}
%
%\begin{claim}\label{claim1}
%\begin{equation}
%\hat{w}_+(\zeta-2/3)=\frac{1}{\sqrt{\pi}}\int_{2/3}^{\zeta}(\zeta-s)^{-1/2}\textcolor{red}{\partial_s}\left(\frac{1}{f'(u)}\right)ds \,\,\qquad s=f(u)
%\end{equation}
%\end{claim}

\begin{lemma}
The following identity holds true
\begin{equation}
{}_2F_1\left(\frac{1}{3},\frac{2}{3};\frac{1}{2};\frac{9}{4}\zeta^2\right)=\frac{1}{1-u^2}\qquad \zeta=\frac{u^3}{3}-u
\end{equation}
\end{lemma}
\begin{proof}From the special case of hypergeometric function (see 15.4.14 DLMF) we have the following identity:
\begin{align*}
{}_2F_1\left(\frac{1}{3},\frac{2}{3};\frac{1}{2};\frac{9}{4}\zeta^2\right)&=%2\cos\left(\frac{1}{3}\arcsin\left(\frac{3}{2}\zeta\right)\right)(4-9\zeta^2)^{-1/2} & \text{Mathematica [Fullsimplify] } \\
\frac{\cos(y)}{\cos(3y)} & 3y=\arcsin\left(\frac{3}{2}\zeta\right)\\
&=\frac{\cos(y)}{\cos(2y)\cos(y)-\sin(2y)\sin(y)} & \\
&=\frac{1}{\cos(2y)-2\sin^2(y)} & \\
&=\frac{1}{1-4\sin^2(y)} & \zeta=2\sin(y)-\frac{8}{3}\sin^3(y)
\end{align*}
Therefore, if $u\defeq -2\sin(y)$, we have $\zeta=\frac{u^3}{3}-u=f(u)$ and \[{}_2F_1\left(\frac{1}{3},\frac{2}{3};\frac{1}{2};\frac{9}{4}\zeta^2\right)=\frac{1}{1-u^2}=-\frac{1}{f'(u)}\]
\end{proof}

%Hence claim \eqref{claim1} is equivalent to
Then equations \eqref{formula1} is equivalent to

\begin{claim}\label{claim 2}
\begin{equation}
\hat{w}_+(\zeta-2/3)=-\frac{1}{\sqrt{\pi}}\int_{2/3}^{\zeta}(\zeta-s)^{-1/2}\textcolor{red}{\partial_s}\left[{}_2F_1\left(\frac{1}{3},\frac{2}{3};\frac{1}{2};\frac{9}{4}s^2\right)\right]ds \qquad \zeta\in(2/3,+\infty)
\end{equation}
\end{claim}

Let us study the RHS of claim \eqref{claim 2}
\begin{align*}
&\int_{2/3}^{\zeta}(\zeta-s)^{-1/2}\textcolor{red}{\partial_s}\left[{}_2F_1\left(\frac{1}{3},\frac{2}{3};\frac{1}{2};\frac{9}{4}s^2\right)\right]ds=2\int_{2/3}^{\zeta}(\zeta-s)^{-1/2}s\,\,{}_2F_1\left(\frac{4}{3},\frac{5}{3};\frac{3}{2};\frac{9}{4}s^2\right)ds & \\
&= -\frac{2}{9}\int_{2/3}^{\zeta}(\zeta-s)^{-1/2}{}_2F_1\left(\frac{5}{3},\frac{7}{3};\frac{5}{2};\frac{1}{2}-\frac{3s}{4}\right)ds+\frac{2}{9}\int_{2/3}^{\zeta}(\zeta-s)^{-1/2}{}_2F_1\left(\frac{5}{3},\frac{7}{3};\frac{5}{2};\frac{1}{2}+\frac{3s}{4}\right)ds & 15.8.28 \text{ DLMF} \\
&=\frac{4}{9\sqrt{3}}\int_0^x(x-y)^{-1/2}\left[{}_2F_1\left(\frac{5}{3},\frac{7}{3};\frac{5}{2};1+y\right)-{}_2F_1\left(\frac{5}{3},\frac{7}{3};\frac{5}{2};y\right)\right]dy & x\in (-\infty,0) \\
&=-\frac{\sqrt{3\pi}}{4}\Gamma(1/6)e^{-\frac{3}{2}\pi i}\int_0^x(x-y)^{-1/2}y^{-5/3}{}_2F_1\left(\frac{5}{3},\frac{1}{6};\frac{1}{3};\frac{1}{y}\right)dy & \\
&=-\frac{\sqrt{3}\pi}{4}\frac{\Gamma(1/6)\Gamma(7/6)}{\Gamma(5/3)}e^{-\frac{3}{2}\pi i} |x|^{-7/6}{}_2F_1\left(\frac{7}{6},\frac{1}{6};\frac{1}{3};\frac{1}{x}\right) & \eqref{4.8} \\
&=\frac{5\pi^2}{24\sqrt{3}}\frac{\Gamma(1/3)}{\Gamma(5/3)}\left({}_2F_1\left(\frac{7}{6},\frac{11}{6};2;\frac{1}{2}-\frac{\zeta}{4}\right)-e^{-\frac{11}{6}\pi i}{}_2F_1\left(\frac{7}{6},\frac{11}{6};2;\frac{1}{2}+\frac{\zeta}{4}\right)\right) & \zeta\in (2/3,+\infty)
%&=-\frac{4i}{9\sqrt{3}}\int_{1}^{\zeta'}(\zeta'-t)^{-1/2} {}_2F_1\left(\frac{5}{3},\frac{7}{3};\frac{5}{2};t\right)dt+\frac{4i}{9\sqrt{3}}\int_{1}^{\zeta'}(\zeta'-t)^{-1/2} {}_2F_1\left(\frac{5}{3},\frac{7}{3};\frac{5}{2};1-t\right)dt & \zeta'=\frac{1}{2}-\frac{3}{4}\zeta\\
%&=\frac{2i}{9\sqrt{3}}\int_{1}^{\zeta'}(\zeta'-t)^{-1/2}{}_2F_1\left(\frac{5}{3},\frac{7}{3};\frac{5}{2};1-t\right)dt -\frac{9i}{32}\textcolor{gray}{\int_{1}^{\zeta'}(\zeta'-t)^{-1/2}{}_2F_1\left(\frac{5}{3},\frac{7}{3};-\frac{1}{2};1-t\right)dt}& 15.10.21 \text{ DLMF }\\
%&=\frac{2i}{9\sqrt{3}}\int_{1}^{\zeta'}(\zeta'-t)^{-1/2} {}_2F_1\left(\frac{1}{6},\frac{5}{6};\frac{5}{2};1-t\right)t^{-3/2}dt &  15.10.13 \text{ DLMF }
\end{align*}
however, ${}_2F_1\left(\frac{7}{6},\frac{11}{6};2;\frac{1}{2}+\frac{\zeta}{4}\right)$ has a branch cut at $\zeta\in(2/3,+\infty)$, thus the claim holds true.

Analougusly, it can be verified for $\hat{w}_-(\zeta+2/3)$ for $\zeta\in(-\infty,-2/3)$.  
%By definition ${}_2F_1\left(\frac{2}{3},\frac{4}{3};\frac{3}{2};t\right)=\sum_{n\geq 0}\frac{(2/3)_n(4/3)_n}{(3/2)_n n!}t^n$, thus
%
%\begin{align*}
%\int_{0}^{\zeta''}(\zeta''-t)^{-1/2} {}_2F_1\left(\frac{2}{3},\frac{4}{3};\frac{3}{2};t\right)dt=\sqrt{\pi}\sum_{n\geq 0}\frac{(2/3)_n(4/3)_n}{(3/2)_n \Gamma(3/2+n)}{\zeta''}^{n+\frac{1}{2}}= 2\sqrt{\zeta''}{}_3F_2\left(\frac{2}{3},\frac{4}{3},1;\frac{3}{2},\frac{3}{2};\zeta''\right)\\
%\int_{1}^{\zeta'}(\zeta'-t)^{-1/2} {}_2F_1\left(\frac{2}{3},\frac{4}{3};\frac{3}{2};t\right)dt= 2\sqrt{\zeta'}{}_3F_2\left(\frac{2}{3},\frac{4}{3},1;\frac{3}{2},\frac{3}{2};\zeta'\right)-\int_0^1(\zeta'-t)^{-1/2} {}_2F_1\left(\frac{2}{3},\frac{4}{3};\frac{3}{2};t\right)dt
%\end{align*}

%Using properties of Caputo's $1/2$-derivative we would like to express $\hat{w}_{\pm}(\zeta)$ as the $1/2$-derivatives of other hypergeometric series. From the integral representation of hypergeomtric functions (see DLMF 15.6.1)
%
%\begin{align*}
%{}_2F_1\left(\frac{7}{6},\frac{11}{6},2,-\frac{3}{4}\zeta\right)&=\frac{3}{5\pi}\int_0^1t^{5/6}(1-t)^{-5/6}(1+\frac{3}{4}\zeta t)^{-7/6}dt & |u|<\frac{4}{3}\\
%&=\frac{3}{5\pi}\zeta^{-1}\int_0^{\zeta}u^{5/6}(\zeta-u)^{-5/6}(1+\frac{3}{4}u)^{-7/6}du &u=\zeta t\\
%&=\frac{3}{5\pi}\int_0^{\zeta}(\zeta-u)^{-1/2}\left(\zeta^{-1/2}(1-\frac{u}{\zeta})^{-1/3}\left(\frac{u}{\zeta}\right)^{5/6}(1+\frac{3}{4}u)^{-7/6}\right)du & \\
%&=\frac{3}{5\sqrt{\pi}}\frac{1}{\Gamma(1/2)}\int_0^{\zeta}(\zeta-u)^{-1/2}\left(\zeta^{-1/2}(1-\frac{u}{\zeta})^{-1/3}\left(\frac{u}{\zeta}\right)^{5/6}(1+\frac{3}{4}u)^{-7/6}\right)du & 
%%&=\textcolor{blue}{\frac{3}{5\sqrt{\pi}}\frac{1}{\Gamma(1/2)}\int_0^{\zeta}(\zeta-u)^{-1/2}\frac{1}{f'(f^{-1}(u))} du } & \\
%%&=\textcolor{blue}{\frac{3}{5\sqrt{\pi}}\zeta^{-1}\partial_{\zeta}^{1/2}f^{-1}(\zeta)} &  
%\end{align*}
%For formula \eqref{formula1} to hold true we must have
%\begin{equation}\label{claim}
%\frac{3}{5\sqrt{\pi}}\left(\zeta^{-1/2}(1-\frac{u}{\zeta})^{-1/3}\left(\frac{u}{\zeta}\right)^{5/6}(1+\frac{3}{4}u)^{-7/6}\right)= \frac{1}{f'(f^{-1}(u))}
%\end{equation}  
%however, $\frac{1}{f'(f^{-1}(u))}$ does not depend on $\zeta$, thus \eqref{claim} is false, and the Airy exponential integral is a counter example of the formula \eqref{formula1}.   
%%thus, we would like to see $\int_0^u(\zeta-t)^{-1/3}t^{5/6}(1+\frac{3}{4}t)^{-7/6}dt$ as the product of first derivatives of an hypergeomtric function and un unknown function $g(\zeta)$
%%\begin{align*}
%%\int_0^u(\zeta-t)^{-1/3}t^{5/6}(1+\frac{3}{4}t)^{-7/6}&\textcolor{blue}{=}\int
%%\end{align*}
%
%\subsection{Second temptative formula} Let us assume that the correct formula that replaces \eqref{formula1} in Theorem \ref{thm:maxim} is 
%\begin{multline}\label{formula2}
%\hat{I}_{\alpha}(\zeta)=\partial^{N/2}_{\zeta, \text{based at }\zeta_\alpha}\left(\int_{f^{-1}(\zeta_\alpha)}^{f^{-1}(\zeta)}\nu\right)=\Gamma(-N/2)\int_{\zeta_\alpha}^\zeta (\zeta-\zeta')^{-N/2}\int_{f^{-1}(\zeta')}\frac{\nu}{df} d\zeta'
%\end{multline} 
%then we compute $\hat{I}$ using properties of the Borel transform of a product:
%
%\begin{align*}
%\hat{I}(\zeta)&=\mathcal{B}(z^{-1/2})\ast\hat{u}(\zeta)\\
%&=\frac{1}{\Gamma(1/2)}\int_0^{\zeta}(\zeta-s)^{-1/2}\hat{u}(s)ds\\
%&=\frac{1}{\Gamma(1/2)}\int_0^\zeta (\zeta-s)^{-1/2}\left(T_{-2/3}\hat{w}_+(s)+T_{2/3}\hat{w}_-(s)\right)ds
%\end{align*}
%
%\begin{align*}
%&=\frac{1}{\Gamma(1/2)}\int_0^\zeta(\zeta-s)^{-1/2} \hat{w}_+(s-2/3)+\frac{1}{\Gamma(1/2)}\int_0^\zeta (\zeta-s)^{-1/2}\hat{w}_-(s+2/3)ds\\
%&=\frac{1}{\Gamma(1/2)}\int_{-2/3}^{\zeta-2/3}(\zeta-\frac{2}{3}-s)^{-1/2} \hat{w}_+(s)+\frac{1}{\Gamma(1/2)}\int_{2/3}^{\zeta+2/3} (\zeta+\frac{2}{3}-s)^{-1/2}\hat{w}_-(s)ds\\
%&=\frac{C_1}{\Gamma(1/2)}\int_{-2/3}^{\zeta-2/3}(\zeta-\frac{2}{3}-s)^{-1/2} {}_2F_1\left(\frac{7}{6},\frac{11}{6},2,-\frac{3}{4}s\right)ds+\frac{C_2}{\Gamma(1/2)}\int_{2/3}^{\zeta+2/3} (\zeta+\frac{2}{3}-s)^{-1/2}{}_2F_1\left(\frac{7}{6},\frac{11}{6},2,\frac{3}{4}s\right)ds\\
%&=\frac{C_3}{\Gamma(1/2)}\int_{-2/3}^{\zeta-2/3}(\zeta-\frac{2}{3}-s)^{-\frac{1}{2}} \frac{\partial}{\partial s}{}_2F_1\left(\frac{1}{6},\frac{5}{6},1,-\frac{3}{4}s\right)ds+\frac{C_4}{\Gamma(1/2)}\int_{2/3}^{\zeta+2/3} (\zeta+\frac{2}{3}-s)^{-\frac{1}{2}}\frac{\partial}{\partial s}{}_2F_1\left(\frac{1}{6},\frac{5}{6},1,\frac{3}{4}s\right)ds\\
%&=\textcolor{red}{C_3\partial^{1/2}_{\zeta, \text{based at } -2/3}{}_2F_1\left(\frac{1}{6},\frac{5}{6},1,-\frac{3}{4}\zeta\right)+C_4\partial^{1/2}_{\zeta, \text{based at } 2/3}{}_2F_1\left(\frac{1}{6},\frac{5}{6},1,\frac{3}{4}\zeta\right)}\, \textcolor{red}{\text{ is wrong!}}
%\end{align*}
%However, ${}_2F_1\left(\frac{1}{6},\frac{5}{6},1,\mp\frac{3}{4}\zeta\right)$ are not algebraic, hence they can not be equal to $\frac{1}{f'(f^{-1}(\zeta))}$ as it is in in Theorem \ref{thm:maxim} part II \eqref{formula2}. 
%
%%Notice that $\pm\frac{2}{3}=f(\mp 1)$, i.e. they are the critical value of $f$. Thus, we  are left to check that ${}_{1}F_{2}\left(\frac{2}{3},\frac{4}{3},\frac{3}{2},\mp\frac{3}{4}\zeta\right) $ is equal to $1/f'(f^{-1}(\zeta)$. Using Mathamtica we simplify the hypergeometric function 
%%
%%\begin{equation}
%%{}_{1}F_{2}\left(\frac{2}{3},\frac{4}{3},\frac{3}{2},-\frac{3}{4}\zeta\right)= \frac{2\sqrt{3}\sinh\left(\frac{2}{3}\mathrm{ArcCsch}\left(\frac{2}{\sqrt{3\zeta}}\right)\right)}{\sqrt{\zeta}\sqrt{4+3\zeta}}
%%\end{equation}
%%
%%then we did the following manipulations: we set $x=\mathrm{ArcCsch}\left(\frac{2}{\sqrt{3\zeta}}\right)$, then  
%%
%%\begin{multline}
%%\frac{2\sqrt{3}\sinh\left(\frac{2}{3}\mathrm{ArcCsch}\left(\frac{2}{\sqrt{3\zeta}}\right)\right)}{\sqrt{\zeta}\sqrt{4+3\zeta}}= \frac{3\sinh\left(\frac{2}{3}x\right)}{2\sinh(x)\sqrt{1+\sinh(x)^2}} \\
%%= \frac{3\sinh\left(\frac{2}{3}x\right)}{2\sinh(x)\cosh(x)}=\frac{3\sinh\left(\frac{2}{3}x\right)}{\sinh(2x)}=\frac{3\sinh\left(\frac{2}{3}x\right)}{3\sinh\left(\frac{2}{3}x\right)+4\sinh^3\left(\frac{2}{3}x\right)}\\
%%=\frac{1}{1+\frac{4}{3}\sinh^2\left(\frac{2}{3}x\right)}
%%\end{multline}
%%
%%With respect to the other hypergeometric function, Mathematica gives
%%
%%\begin{equation}
%%{}_{1}F_{2}\left(\frac{2}{3},\frac{4}{3},\frac{3}{2},\frac{3}{4}\zeta\right)= \frac{2\sqrt{3}\sin\left(\frac{2}{3}\mathrm{ArcCsc}\left(\frac{2}{\sqrt{3\zeta}}\right)\right)}{\sqrt{\zeta}\sqrt{4-3\zeta}}
%%\end{equation}
%% now we set $x=\frac{2}{3}\mathrm{ArcCsc}\left(\frac{2}{\sqrt{3\zeta}}\right)$, then 
%% \begin{multline}
%% \frac{2\sqrt{3}\sin\left(\frac{2}{3}\mathrm{ArcCsc}\left(\frac{2}{\sqrt{3\zeta}}\right)\right)}{\sqrt{\zeta}\sqrt{4-3\zeta}}=\frac{2\sqrt{3}\sin(x)}{\frac{2}{\sqrt{3}}\sin\left(\frac{3}{2}x\right)\sqrt{4-4\sin^2\left(\frac{3}{2}x\right)}}\\
%% \frac{3\sin(x)}{2\sin\left(\frac{3}{2}x\right)\cos\left(\frac{3}{2}x\right)}=\frac{3\sin(x)}{\sin(3x)}=\frac{3\sin(x)}{3\sin(x)-4\sin^3(x)}\\
%% =\frac{1}{1-\frac{4}{3}\sin^2(x)}
%% \end{multline}
%
%
%%\begin{remark}
%%Notice that the half derivative of a simple pole gives a log singularity; indeed 
%%\begin{align*}
%%\frac{1}{1-\frac{4}{3}\sin^2(x)} \text{has simple poles at } \zeta=\frac{4}{3} \text{ and } \zeta=\infty \\
%%\frac{1}{1+\frac{4}{3}\sinh^2(x)} \text{has simple poles at } \zeta=\frac{4}{3} \text{ and } \zeta=\infty \\ 
%%\end{align*} 
%%\end{remark}
%%
%%
%%If $\tilde{I}(z)$ is Borel--Laplace summable, then for a suitable choice of $U_1,U_2$ and of a direction $\theta\in\R/2\pi i\Z$, $\mathcal{S}^{\theta}\tilde{I}(z)=I(z)$. Indeed 
%%\begin{align*}
%%I(z)&=\int_{\mathcal{C}} e^{-zf(t)}\text{dt}=\int_{\mathcal{C}_h} e^{-z\zeta}\frac{d\zeta}{f'(f^{-1}(\zeta))}=e^{-2/3z}\int_{T_{2/3}\mathcal{C}_h}e^{-z\zeta}\frac{d\zeta}{f'(f^{-1}(\zeta+2/3))}. 
%%\end{align*}
%%
%%However we are interested in resurgence properties of $\tilde{I}(z)$, thus we take its Borel transform. 
%\subsubsection{Comparison with Aaron}
%
%The Airy integral can be written in terms of the modified Bessel equation as  
%
%\begin{equation}
%Ai(x)=\frac{1}{\pi\sqrt{3}}x^{1/2}K(\frac{2}{3}x^{3/2}).
%\end{equation}
%On the other hand we have, for $z=x^{3/2}$
%\begin{equation}
%Ai(x)=-\frac{z^{1/3}}{2\pi i}I(z)=-\frac{1}{2\pi i}x^{1/2}I(x^{3/2}) 
%\end{equation} 
%hence
%\begin{align*}
%-\frac{1}{2\pi i}I(x^{3/2})&=\frac{1}{\pi\sqrt{3}}K(\frac{2}{3}x^{3/2})\\
%\frac{i}{2} I(z)&=\frac{1}{\sqrt{3}}K(\frac{2}{3}z)
%\end{align*}
%In particular, the Borel trasforms of LHS and RHS\footnote{The conjugate variable of $z$ is $\zeta$, hence $\hat{K}(\frac{2}{3}z)=\frac{3}{2}\hat{K}(\frac{3}{2}\zeta)$. Indeed assuming $K(z)=\sum_{n\geq 0}a_nz^{-n}$, the Borel transform of $K(\frac{2}{3}z)=\sum_{n\geq 0}a_n\left(\frac{3}{2}\right)^nz^{-n}$ is by definition \[\hat{K}(\frac{2}{3}z)=\sum_{k\geq 0}a_{n+1}\left(\frac{3}{2}\right)^{n+1}\frac{\zeta^n}{n!}=\frac{3}{2}\sum_{k\geq 0}\frac{a_{n+1}}{n!}\left(\frac{3}{2}\zeta\right)^{n}=\frac{3}{2}\hat{K}(\frac{3}{2}\zeta). \]} must be equal, i.e.
%\begin{align*}
%\frac{i}{2} \hat{I}(\zeta)&=\frac{\sqrt{3}}{2}\hat{K}(\frac{3}{2}\zeta)\\
%&= \frac{\sqrt{3}}{2}\frac{1}{\sqrt{2}}(\frac{3}{2}\zeta-1)^{-1/2}{}_2F_1\left(\frac{1}{6},\frac{5}{6};\frac{1}{2};\frac{1}{2}-\frac{3}{4}\zeta\right)\\
%&=\frac{\sqrt{3}}{2}(3\zeta-2)^{-1/2}{}_2F_1\left(\frac{1}{6},\frac{5}{6};\frac{1}{2};\frac{1}{2}-\frac{3}{4}\zeta\right)\\
%&=\frac{\sqrt{3}}{2}T_{-2/3}\left((3\zeta)^{-1/2}{}_2F_1\left(\frac{1}{6},\frac{5}{6};\frac{1}{2};-\frac{3}{4}\zeta\right)\right)\\
%&=\frac{1}{2}T_{-2/3}\left(\frac{1}{\sqrt{\zeta}}{}_2F_1\left(\frac{1}{6},\frac{5}{6};\frac{1}{2};-\frac{3}{4}\zeta\right)\right)\\
%\hat{I}(\zeta)&=-i T_{-2/3}\left(\frac{1}{\sqrt{\zeta}}\,\,{}_2F_1\left(\frac{1}{6},\frac{5}{6};\frac{1}{2};-\frac{3}{4}\zeta\right)\right)
%\end{align*}

%%In addition, since $\hat{I}(\zeta)=C_3\partial_{\zeta,\text{based at } -2/3}^{1/2} {}_2F_1\left(\frac{1}{6},\frac{5}{6};\frac{1}{2};-\frac{3}{4}\zeta\right)$ is a solution of the Borel transform of equation \eqref{eq:I}; thus
%%\begin{align*}
%%C_3\partial_{\zeta,\text{based at } -2/3}^{1/2} {}_2F_1\left(\frac{1}{6},\frac{5}{6};\frac{1}{2};-\frac{3}{4}\zeta\right)&=-i T_{-2/3}\left(\frac{1}{\sqrt{\zeta}}\,\,{}_2F_1\left(\frac{1}{6},\frac{5}{6};\frac{1}{2};-\frac{3}{4}\zeta\right)\right)\\
%%C_3\partial_{\zeta}^{1/2} {}_2F_1\left(\frac{1}{6},\frac{5}{6};\frac{1}{2};-\frac{3}{4}\zeta\right)&=-i \frac{1}{\sqrt{\zeta}}\,\,{}_2F_1\left(\frac{1}{6},\frac{5}{6};\frac{1}{2};-\frac{3}{4}\zeta\right)
%%\end{align*} 
%
%Equation 5 in Aaron is 
%\begin{equation}
%\left[\frac{\partial^2}{\partial z^2}+2\frac{\partial}{\partial z}+\frac{1}{z}\frac{\partial}{\partial z}+\frac{1}{z}-\frac{1}{9z^2}\right]\kappa=0
%\end{equation}
%Its Borel transform is 
%\begin{align*}
%&\zeta^2\hat{\kappa}-2\zeta\hat{\kappa}+1\ast(-\zeta\hat{\kappa}+\hat{\kappa})-\frac{1}{9}\zeta\ast\hat{\kappa}=0 \\
%&\zeta^2\hat{\kappa}-2\zeta\hat{\kappa}-\int_{0}^\zeta (s\hat{\kappa}(s)-\hat{\kappa}(s))ds-\frac{1}{9}\int_0^{\zeta}(\zeta-s)\hat{\kappa}(s)ds=0 \\
%&\, \text{ taking derivatives once} \\
%& 2\zeta\hat{\kappa}+\zeta^2\hat{\kappa}'-2\hat{\kappa}-2\zeta\hat{\kappa}'-\zeta\hat{\kappa}+\hat{\kappa}-\frac{1}{9}\int_0^{\zeta}\hat{k}(s)ds=0\\
%& \zeta\hat{\kappa}+\zeta^2\hat{\kappa}'-\hat{\kappa}-2\zeta\hat{\kappa}'-\frac{1}{9}\int_0^{\zeta}\hat{k}(s)ds=0\\
%&\, \text{ taking derivatives once again} \\
%& \hat{\kappa}+\zeta\hat{\kappa}'+2\zeta\hat{k}'+\zeta^2\hat{\kappa}''-\hat{\kappa}'-2\hat{\kappa}'-2\zeta\hat{\kappa}''-\frac{1}{9}\hat{\kappa}=0\\
%&(\zeta^2-2\zeta)\hat{\kappa}''+(3\zeta-3)\hat{k}'+\frac{8}{9}\hat{\kappa}=0
%\end{align*} 
%The last equation is a hypergeomtric equation and the fundamental solutions are (see DLMF 15.10.2) 
%\begin{equation}
%\hat{\kappa}_1(\zeta)={}_2F_1\left(\frac{2}{3},\frac{4}{3};\frac{3}{2};\frac{\zeta}{2}\right)
%\end{equation}
%\begin{equation}
%\hat{\kappa}_2(\zeta)=\frac{1}{\sqrt{2}}\zeta^{-1/2}{}_2F_1\left(\frac{1}{6},\frac{5}{6};\frac{1}{2};\frac{\zeta}{2}\right)
%\end{equation}
%
%In particular, comparing with $\hat{I}(\zeta)$ we have 
%
%\begin{equation}
%\hat{I}\left(\frac{2}{3}\zeta\right)=-i\sqrt{3}\hat{\kappa}_2(\zeta-1)
%\end{equation}
%
%------------------------------------------------------------------------
%
%However, at pag. 5 in Aaron there is a different solution, namely 
%\begin{align*}
%\hat{\kappa}(\zeta-1)&=-i\sqrt{2}(\zeta-1)^{-1/2}{}_2F_1\left(\frac{1}{6},\frac{5}{6};\frac{1}{2};\frac{1}{2}-\frac{\zeta}{2}\right)\\
%\hat{\kappa}(\zeta)&=-i\sqrt{2}\zeta^{-1/2}{}_2F_1\left(\frac{1}{6},\frac{5}{6};\frac{1}{2};-\frac{\zeta}{2}\right)=-2i\frac{\cosh\left(\frac{2}{3}\mathrm{arcsinh}\left(\frac{\sqrt{\zeta}}{\sqrt{2}}\right)\right)}{\sqrt{\zeta}\sqrt{2+\zeta}}
%\end{align*} 
%------------------------------------------------------------------------
%
%
%As written by Aaron in Section 2.3, the asymptotic behvaior of $\kappa(z)$ is given by 
%\begin{equation}
%\kappa(z)\sim \left(\frac{\pi}{2z}\right)^{1/2}\sum_{n\geq 0}a_n(1/3)z^{-n}
%\end{equation}
%where $a_n(1/3)=\frac{(1/6)_n(5/6)_n}{(-2)^nn!}$. Hence the Borel transform of $\kappa(z)$ is
%\begin{align*}
%\hat{\kappa}(\zeta)&=\left(\frac{\pi}{2}\right)^{1/2}\sum_{n\geq 0}a_n(1/3)\frac{\zeta^{n-1/2}}{\Gamma(n+1/2)}\\
%&=\left(\frac{\pi}{2}\right)^{1/2}\sum_{n\geq 0}\frac{(1/6)_n(5/6)_n}{(-2)^nn!}\frac{\zeta^{n-1/2}}{\Gamma(n+1/2)}\\
%&=\frac{1}{\sqrt{2}}\zeta^{-1/2}\sum_{n\geq 0}\frac{(1/6)_n(5/6)_n}{(1/2)_nn!}\left(-\frac{\zeta}{2}\right)^n\\
%&=\frac{1}{\sqrt{2}}\zeta^{-1/2}{}_2F_1\left(\frac{1}{6}\frac{5}{6};\frac{1}{2};-\frac{\zeta}{2}\right)
%\end{align*}
%\begin{align*}
%\hat{\kappa}(\zeta)&=\left(\frac{\pi}{2}\right)^{1/2}\mathcal{B}(z^{-1/2})\ast\left(\delta+\sum_{n\geq 0}a_{n+1}(1/3)\frac{\zeta^n}{n!}\right)=\frac{1}{\sqrt{2}}\zeta^{-1/2}\ast\left(\delta-\frac{5}{72}{}_2F_1\left(\frac{7}{6},\frac{11}{6};2;-\frac{\zeta}{2}\right)\right)\\
%&=\frac{1}{\sqrt{2\zeta}}-\frac{5}{72\sqrt{2}}\int_0^{\zeta}(\zeta-s)^{-1/2}{}_2F_1\left(\frac{7}{6},\frac{11}{6};2;-\frac{s}{2}\right)ds\\
%&=\frac{1}{\sqrt{2\zeta}}+\frac{1}{\sqrt{2}}\int_0^{\zeta}(\zeta-s)^{-1/2}\frac{d}{ds}\left({}_2F_1\left(\frac{1}{6},\frac{5}{6};1;-\frac{s}{2}\right)\right)ds\\
%&\textcolor{red}{=\frac{1}{\sqrt{2}}\partial_\zeta^{1/2}{}_2F_1\left(\frac{1}{6},\frac{5}{6};1;-\frac{\zeta}{2}\right)}\, \textcolor{red}{\text{ is wrong!}}\\
%\end{align*}
%
%\subsubsection{Algebraic Hypergeometric}
%Aaron cliams 
%\begin{equation}
%{}_2F_1\left(\frac{1}{3},\frac{2}{3};\frac{1}{2};\zeta^2\right)=\textcolor{red}{-}\frac{1}{4u^2-1}\qquad \text{wtih } \zeta=-4u^3+3u
%\end{equation}
%\begin{proof}
%\begin{align*}
%{}_2F_1\left(\frac{1}{3},\frac{2}{3};\frac{1}{2};\zeta^2\right)&=\cos\left(\frac{1}{3}\arcsin(\zeta)\right)(1-\zeta^2)^{-1/2} & \text{Mathematica [Fullsimplify] } \\
%&=\frac{\cos(y)}{\cos(3y)} & 3y=\arcsin(\zeta)\\
%&=\frac{\cos(y)}{\cos(2y)\cos(y)-\sin(2y)\sin(y)} & \\
%&=\frac{1}{\cos(2y)-2\sin^2(y)} & \\
%&=\frac{1}{1-4\sin^2(y)} & \zeta=3\sin(y)-4\sin^3(y)
%\end{align*}
%Therefore, if $u\defeq\sin(y)$, we have $\zeta=-4u^3+3u$ and \[{}_2F_1\left(\frac{1}{3},\frac{2}{3};\frac{1}{2};\zeta^2\right)=\frac{1}{1-4u^2}=\frac{1}{3}\frac{1}{f'(u)}\]
%\end{proof}
%Now, we notice that 
%\begin{align*}
%{}_2F_1\left(\frac{1}{3},\frac{2}{3};\frac{1}{2};\zeta^2\right)&=2 {}_2F_1\left(\frac{2}{3},\frac{4}{3};\frac{3}{2};\frac{1}{2}-\frac{\zeta}{2}\right)+2 {}_2F_1\left(\frac{2}{3},\frac{4}{3};\frac{3}{2};\frac{1}{2}+\frac{\zeta}{2}\right) & 15.8.27 \text{ DLMF} \\
%&=-{}_2F_1\left(\frac{2}{3},\frac{4}{3};\frac{3}{2};\frac{\zeta}{2}\right)-3\sqrt{3}\zeta^{-1/2}{}_2F_1\left(\frac{1}{6},\frac{5}{6};\frac{1}{2};\frac{\zeta}{2}\right)+ & 15.10.17 \text{DLMF}\\
%&\qquad -{}_2F_1\left(\frac{2}{3},\frac{4}{3};\frac{3}{2};-\frac{\zeta}{2}\right)+3\sqrt{3}i\zeta^{-1/2}{}_2F_1\left(\frac{1}{6},\frac{5}{6};\frac{1}{2};-\frac{\zeta}{2}\right) & 15.10.17 \text{DLMF}\\
%&=-\hat{\kappa}_1(\zeta)-3\sqrt{6}\hat{\kappa}_2(\zeta)-\hat{\kappa}_1(-\zeta)+3\sqrt{6}\hat{\kappa}_2(-\zeta) \\
%&=-(\hat{\kappa}_1(\zeta)+\hat{\kappa}_1(-\zeta))-3\sqrt{6}
%\end{align*}
\end{example}

\begin{example}[Bessel]

Let $X=\C^*$, $f(x)=x+\frac{1}{x}$ and $\nu=\frac{dx}{x}$, then the ciritcal points of $f$ are $x=\pm1$ and 
\begin{equation}
I(z)\defeq\int_0^{\infty}e^{-zf(x)}\frac{dx}{x}.
\end{equation}

By change of cooridnates $t=zx$
\begin{align*}
I(z)=\int_0^{\infty}e^{-z\left(\frac{t}{z}+\frac{z}{t}\right)}\frac{dt}{t}=\int_0^{\infty}e^{-\left(t+\frac{z^2}{t}\right)}\frac{dt}{t}=2K_0(2z) \,\qquad |\arg z|<\frac{\pi}{4}
\end{align*}
where $K_0(z)$ is the modified Bessel function (see definition 10.32.10 DLMF). In particular, since $K_0(z)$ solves 
\begin{equation}
\frac{d^2}{dz^2}w(z)+\frac{1}{z}\frac{d}{dz}w(z)-w(z)=0
\end{equation}
and $K_0(z)\sim\left(\frac{\pi}{2}\right)^{1/2}e^{-z}z^{-1/2}\sum_{k\geq 0}\frac{(1/2)_k(1/2)_k}{(-2)^kk!}z^{-k}$ as $z\to\infty$ (see DLMF 10.40.2), then $I(z)$ is a solution of 
\begin{equation}
\label{eq:IB}
\frac{d^2}{dz^2}I(z)+\frac{1}{z}\frac{d}{dz}I(z)-4I(z)=0.
\end{equation} 
The formal integral of \eqref{eq:IB} is given by a two parameter formal solution $\tilde{I}_1(z)$
\begin{equation}
\tilde{I}(z)=\sum_{\mathbf{k}\in\N^2}U^ke^{-\mathbf{k}\cdot\lambda z}z^{-\tau\cdot \mathbf{k}}\tilde{w}_{\mathbf{k}}(z)
\end{equation}
where $\lambda=(2,-2)$, $\tau=(-\frac{1}{2},-\frac{1}{2})$, $U^k:=U_1^{k_1}U_2^{k_2}$ with $k=(k_1,k_2)$ and $U_1,U_2\in\C$, and $\tilde{w}_{\mathbf{k}}(z)\in\C[[z^{-1}]]$ is a formal solution of 
\begin{multline}
\tilde{w}_{\mathbf{k}}''(z)-4(k_1-k_2)\tilde{w}_{\mathbf{k}}'(z)+4(1-(k_1-k_2)^2)\tilde{w}_{\mathbf{k}}(z)+\frac{(k_1+k_2-1)}{z}\tilde{w}_{\mathbf{k}}'(z)+\\
-2(k_1-k_2)\frac{(k_1+k_2-1)}{z}\tilde{w}_{\mathbf{k}}(z)+\frac{(k_1+k_2)^2}{4z^2}\tilde{w}_{\mathbf{k}}(z)=0
\end{multline}
The only non zero $\tilde{w}_{\mathbf{k}}(z)$ occurs for $\mathbf{k}=(1,0)$ and $\mathbf{k}=(0,1)$, hence 
\begin{equation}
\tilde{I}(z)=U_1 e^{-2z}z^{-1/2}\tilde{w}_{(1,0)}(z)+U_2e^{2z}z^{-1/2}\tilde{w}_{(0,1)}(z)
\end{equation}
and according to our convention, we define
\begin{align}
\label{IB+} &\tilde{I}_{1}(z)\defeq e^{-2z}z^{-1/2}\tilde{w}_{(1,0)}(z)\\
\label{IB-} &\tilde{I}_{-1}(z)\defeq e^{2z}z^{-1/2}\tilde{w}_{(0,1)}(z).
\end{align}
We set $\tilde{w}_{(1,0)}=\tilde{w}_+$ and $\tilde{w}_{(0,1)}=\tilde{w}_-$, then their Borel transforms are solutions respectively of the following equations 
\begin{align*}
(+)\quad &\zeta^2\hat{w}_+(\zeta)+4\zeta\hat{w}_+(\zeta)+\frac{1}{4}\zeta\ast\hat{w}_+(\zeta)=0\\
(-)\quad &\zeta^2\hat{w}_+(\zeta)-4\zeta\hat{w}_+(\zeta)+\frac{1}{4}\zeta\ast\hat{w}_+(\zeta)=0
\end{align*}
taking twice derivative in $\zeta$ we get
\begin{align*}
(+)\quad &(\zeta^2+4\zeta)\frac{d^2}{d\zeta^2}\hat{w}_++4(\zeta-1)\frac{d}{d\zeta}\hat{w}_++\frac{9}{4}\hat{w}_+=0 &\\
(-)\quad &(\zeta^2-4\zeta)\frac{d^2}{d\zeta^2}\hat{w}_-+4(\zeta+1)\frac{d}{d\zeta}\hat{w}_++\frac{9}{4}\hat{w}_-=0 &\\
(+)\quad &\xi(1-\xi)\frac{d^2}{d\xi^2}\hat{w}_++(1-4\xi)\frac{d}{d\xi}\hat{w}_+-\frac{9}{4}\hat{w}_+=0 & \xi=-\frac{\zeta}{4} \\
(-)\quad &\xi(1-\xi)\frac{d^2}{d\xi^2}\hat{w}_-+(1-4\xi)\frac{d}{d\xi}\hat{w}_--\frac{9}{4}\hat{w}_-=0 & \xi=\frac{\zeta}{4}
\end{align*}
therefore, since equation $(+), (-)$ are hypergeometric the fundamental solution is (see DLMF 15.10.2)
\begin{align}
\label{w+}&\hat{w}_+(\zeta)={}_2F_1\left(\frac{3}{2},\frac{3}{2};2;-\frac{\zeta}{4}\right)\\
\label{w+}&\hat{w}_-(\zeta)={}_2F_1\left(\frac{3}{2},\frac{3}{2};2;\frac{\zeta}{4}\right)
\end{align}
In particular, we notice that taking the series expansion of $\hat{w}_+$ and $\hat{w}_-$ we get numerically that
\begin{align*}
\hat{w}_+(\zeta-4)=\frac{1}{\pi}\log(z)\hat{w}_-(z)+\phi_{\text{reg}}\\
\hat{w}_-(\zeta+4)=\frac{1}{\pi}\log(z)\hat{w}_+(z)+\psi_{\text{reg}}
\end{align*}
and analytically (thanks to 15.2.3 DLMF)
\begin{align*}
\hat{w}_+(\zeta+i0)-\hat{w}_+(\zeta-i0)&={}_2F_1\left(\frac{3}{2},\frac{3}{2};2;-\frac{\zeta}{4}+i0\right)-{}_2F_1\left(\frac{3}{2},\frac{3}{2};2;-\frac{\zeta}{4}-i0\right) &\qquad\zeta<-4\\
&=-8i\left(-\frac{\zeta}{4}-1\right)^{-1}\sum_{n\geq 0}\frac{(1/2)_n(1/2)_n}{n!\Gamma(n)}\left(\frac{\zeta}{4}+1\right)^n\\
&=8i\sum_{n\geq 0}\frac{(1/2)_n(1/2)_n}{n!\Gamma(n)}\left(\frac{\zeta}{4}+1\right)^{n-1} & \\
&=8i\sum_{n\geq 0}\frac{(1/2)_{n+1}(1/2)_{n+1}}{(n+1)!\Gamma(n+1)}\left(\frac{\zeta}{4}+1\right)^{n} &\\
&=\mathbf{2i}\sum_{n\geq 0}\frac{(3/2)_{n}(3/2)_{n}}{(n)!\Gamma(n+2)}\left(\frac{\zeta}{4}+1\right)^{n} & \\
&=\mathbf{2i}{}_2F_1\left(\frac{3}{2},\frac{3}{2};2;\frac{\zeta}{4}+1\right)
\end{align*}
\begin{align*}
\hat{w}_-(\zeta+i0)-\hat{w}_-(\zeta-i0)&={}_2F_1\left(\frac{3}{2},\frac{3}{2};2;\frac{\zeta}{4}+i0\right)-{}_2F_1\left(\frac{3}{2},\frac{3}{2};2;\frac{\zeta}{4}-i0\right)& \qquad\zeta>4\\
&=8i\left(\frac{\zeta}{4}-1\right)^{-1}\sum_{n\geq 0}\frac{(1/2)_n(1/2)_n}{n!\Gamma(n)}\left(\frac{\zeta}{4}-1\right)^n\\
&=-8i\sum_{n\geq 0}\frac{(1/2)_n(1/2)_n}{n!\Gamma(n)}\left(1-\frac{\zeta}{4}\right)^{n-1}\\
&=-8i\sum_{n\geq 0}(-1)^n\frac{(1/2)_{n+1}(1/2)_{n+1}}{(n+1)!\Gamma(n+1)}\left(1-\frac{\zeta}{4}\right)^{n}\\
&=\mathbf{-2i}\sum_{n\geq 0}(-1)^n\frac{(3/2)_{n}(3/2)_{n}}{(n)!\Gamma(n+2)}\left(1-\frac{\zeta}{4}\right)^{n}\\
&=\mathbf{-2i}{}_2F_1\left(\frac{3}{2},\frac{3}{2};2;1-\frac{\zeta}{4}\right)
\end{align*}
These are evidence of the resurgent properties of $\tilde{I}_{\pm 1}(z)$.

\begin{lemma}
The following identity holds true
\begin{equation}
{}_2F_1\left(\frac{1}{2},\frac{1}{2};\frac{1}{2};\frac{\zeta^2}{4}\right)=2i\frac{u}{u^2-1}\qquad \qquad \zeta=u+\frac{1}{u}
\end{equation}
\end{lemma}
\begin{proof}
From 15.4.13 DLMF, we have
\begin{align*}
{}_2F_1\left(\frac{1}{2},\frac{1}{2};\frac{1}{2};\frac{\zeta^2}{4}\right)&=\frac{2}{\sqrt{4-\zeta^2}} & y=\mathrm{arccsc}(\zeta/2)\\
&=\frac{1}{\sqrt{1-\mathrm{csc}^2(y)}} & \\
&=-i\tan(y) &   \zeta=\frac{2}{\sin(y)}
\end{align*}
therefore if $u=\tan\left(\frac{y}{2}\right)$, we have $\zeta=\frac{1+u^2}{u}=f(u)$ and 
\begin{equation*}
{}_2F_1\left(\frac{1}{2},\frac{1}{2};\frac{1}{2};\frac{\zeta^2}{4}\right)=2i\frac{u}{u^2-1}=\frac{2i}{f'(u)u}
\end{equation*}  
\end{proof}

\begin{claim}
\label{Bessel}
\begin{equation}
\hat{w}_{+}(\zeta-2)=i\pi\int_{2}^\zeta(\zeta-\zeta')^{-1/2}2\zeta'\,\,{}_2F_1\left(\frac{3}{2},\frac{3}{2};\frac{3}{2};\frac{\zeta^2}{4}\right)d\zeta'\,\qquad \zeta\in(2,+\infty)
\end{equation}
\end{claim}
\begin{proof}
Let us first consider the RHS of \eqref{Bessel}
\begin{align*}
&2\pi\int_{2}^\zeta(\zeta-\zeta')^{-1/2}\zeta'\,\,{}_2F_1\left(\frac{3}{2},\frac{3}{2};\frac{3}{2};\frac{\zeta^2}{4}\right)d\zeta'= &\\
&\quad = \frac{4}{3}\int_{2}^\zeta(\zeta-\zeta')^{-1/2}\left[{}_2F_1\left(2,2;\frac{5}{2};\frac{1}{2}+\frac{\zeta'}{4}\right)-{}_2F_1\left(2,2;\frac{5}{2};\frac{1}{2}-\frac{\zeta'}{4}\right)\right]d\zeta' &\\
&\quad =\frac{8}{3}\int_0^x(y-x)^{-1/2}\left[{}_2F_1\left(2,2,;\frac{5}{2};y\right)-{}_2F_1\left(2,2;\frac{5}{2};1-y\right)\right]dy & x\in (-\infty,0)\\
&\quad =2\pi\int_0^x(x-y)^{-1/2}y^{-2}F\left(2,\frac{1}{2};1;\frac{1}{y}\right)dy & \eqref{4.8} \\
&\quad =\pi^2|x|^{-3/2}{}_2F_1\left(\frac{3}{2},\frac{1}{2};1;\frac{1}{x}\right) & x\in (-\infty, 0)\\
&\quad =\frac{\pi^2}{2}\left({}_2F_1\left(\frac{3}{2},\frac{3}{2};2;\frac{1}{2}-\frac{\zeta}{4}\right)-i{}_2F_1\left(\frac{3}{2},\frac{3}{2};2;\frac{1}{2}+\frac{\zeta}{4}\right)\right) & \zeta\in(2,+\infty)
\end{align*}
however ${}_2F_1\left(\frac{3}{2},\frac{3}{2};2;\frac{1}{2}+\frac{\zeta}{4}\right)$ has a branch cut at $\zeta\in(2,+\infty)$, thus the claim holds true. 
\end{proof}
Analougusly, it can be verified for $\hat{w}_-(\zeta+2)$ for $\zeta\in(-\infty,-2)$.  
\end{example}

\section{Useful identities for Gauss hypergeomtric functions}

\begin{multline}
\label{w1w3w5}
{}_2F_1\left(a,b;c;z\right)=e^{b\pi i}\frac{\Gamma(c)\Gamma(a-c+1)}{\Gamma(a+b-c+1)\Gamma(c-b)}{}_2F_1\left(a,b;c;1-z\right)+\\
-e^{(a+b-c)\pi i}\frac{\Gamma(c)\Gamma(a-c+1)}{\Gamma(b)\Gamma(a-b+1)}|z|^{-a}{}_2F_1\left(a,a-c+1;a-b+1;\frac{1}{z}\right)
\end{multline}

\begin{multline}
\label{4.5}
\int_0^x |y|^{a-\mu-1}{}_2F_1\left(a,b;c,y\right)(x-y)^{\mu-1}dy=\frac{\Gamma(\mu)\Gamma(a-\mu)}{\Gamma(a)}|x|^{a-1}{}_2F_1\left(a-\mu,b;c;x\right)\\
x\in(-\infty,0)\cup(0,1), \, \Re a>\Re\mu>0
\end{multline}
which can be rewritten as 
\begin{multline}
\label{4.8}
\int_{y>x}|y|^{-a}|x-y|^{\mu-1}{}_2F_1\left(a,b;c;y^{-1}\right)dy=\frac{\Gamma(\mu)\Gamma(a-\mu)}{\Gamma(a)}|x|^{-a+\mu}{}_2F_1\left(a-\mu,b;c;x^{-1}\right)\\
x\in(-\infty,0)\cup(1,\infty), \, \Re a>\Re\mu>0
\end{multline}
%\textcolor{red}{We should check that for Hypergeomtric functions the following relation holds true
%\[{}_2F_1(a,b;c;z-1)\propto\int_1^z(z-t)^{-1/2}\left[{}_2F_1\left(a+\frac{1}{2},b+\frac{1}{2};c+\frac{1}{2};1-t\right)-{}_2F_1\left(a+\frac{1}{2},b+\frac{1}{2};c+\frac{1}{2};t\right)\right]dt\]
%}
%
%The following identities hold true
%\begin{multline}
%\label{a+b+c+}
%\int_{0<\frac{y}{x}<1}|y|^{c-1}(1-y)^{a+b-c}{}_2F_1\left(a,b;c;y\right)\frac{|x-y|^{\mu-1}}{\Gamma(\mu)}dy=\\
%\frac{\Gamma(c)}{\Gamma(c+\mu)}|x|^{c-1+\mu}(1-x)^{a+b-c+\mu}{}_2F_1\left(a+\mu,b+\mu;c+\mu;x\right)\\
% x\in (-\infty,0)\cup(0,1), \,\Re c>0,\,\Re\mu>0
%\end{multline}
%
%\begin{multline}\label{a-b-c-}
%\int_{-\infty}^x{}_2F_1\left(a,b;c;y\right) (x-y)^{\mu-1}dy=\Gamma(\mu)\frac{\Gamma(a-\mu)}{\Gamma(a)}\frac{\Gamma(b-\mu)}{\Gamma(b)}\frac{\Gamma(c)}{\Gamma(c-\mu)}{}_2F_1\left(a-\mu,b-\mu;c-\mu;x\right)\\
%x<1, \Re a,\Re b>\Re\mu>0
%\end{multline}

\section{Resurgence for degree $3$ polynomials}

Let $f$ be a degree $3$ polynomial, and $t_{1},t_{2}$ its critical points (not necessarily distinguished):
\begin{enumerate}
\item if $t_{1}\neq t_{2}$, then 
\[I(z)=\int_{\mathcal{C}_{j}}e^{-zf}dt\]
is a solution of 
\begin{equation}
\label{eq:I for p not 0}
 I''+a I'+bI+c\frac{I'}{z}+\frac{d}{z}I+\frac{e}{z^2}I=0
\end{equation}
where $a,b,c,d,e$ are determined in terms of $f$.
\item if $t_{1}=t_2$, then 
\[I(z)=\int_{\mathcal{C}_{1}}e^{-zf}dt\]
is a solution of a first order ODE
\begin{equation}
I'+\left(a_4-\frac{a_2^3}{27a_1^2}+\frac{1}{3z}\right)I=0
\end{equation}
\end{enumerate}

\begin{proof}
Let $f(t)=a_1t^3+a_2t^2+a_3t+a_4$ with $a_1\neq 0$, 
\begin{align*}
\int_{\mathcal{C}_j}e^{-fz}dt&=\int_{\mathcal{C}_j+\frac{a_2}{3a_1}}e^{-(a_1t^3+pt+q)z}dt & t\to t-\frac{a_2}{3a_1}
\end{align*}
where $p=a_3-\frac{a_2^2}{3a_1}$ and $q=a_4-\frac{a_2a_3}{3a_1}+\frac{2a_2^3}{27a_1^2}$.

Case $(1)$: if $p\neq 0$,
\begin{align*}
I(z)=\int t(3a_1t^2+p)ze^{-fz}=\int(3a_1t^3+pt)ze^{-ft}=\\
\int2a_1t^3ze^{-fz}+\int(a_1t^3+pt+q)ze^{-fz}-qzI\\
2z\int a_1t^3e^{-fz}-zI'-qzI
\end{align*}

\begin{align*}
2z\int a_1t^3e^{-fz}=2z^2\int\frac{t^4}{4}a_1(3a_1t^2+p)e^{-fz}=\frac{z^2}{2}\int (3a_1^2t^6+pa_1t^4)e^{-fz}=\\
\frac{z^2}{2}\int(3a_1^2t^6+6pa_1t^4+3q^2+3p^2t^2+6pqt+6a_1qt^3)e^{-fz}+\\
+\frac{z^2}{2}\int(pa_1t^4-6pa_1t^4)e^{-fz}-\frac{z^2}{2}\int(3q^2+3p^2t^2+6pqt+6a_1qt^3)e^{-fz}\\
=\frac{3z^2}{2}I''+\frac{3z^2}{2}q^2I+3qz^2I'-\frac{z^2}{2}p\int(3a_1t^4+pt^2)e^{-fz}-z^2p\int(a_1t^4+pt^2)e^{-fz}\\
=\frac{3z^2}{2}I''+\frac{3z^2}{2}q^2I+3qz^2I'-\frac{5}{3}zp\int te^{-fz}-\frac{2}{3}z^2p^2\int t^2e^{-fz}
\end{align*}

hence

\begin{align}
I=-zI'-qzI+\frac{3z^2}{2}I''+\frac{3z^2}{2}q^2I+3qz^2I'-\frac{5}{3}zp\int te^{-fz}-\frac{2}{3}z^2p^2\int t^2e^{-fz}\\
\frac{3z^2}{2}\left(I''+q^2I+2qI'-\frac{2}{3z}I'-\frac{2q}{3z}I-\frac{2}{3z^2}I-\frac{10}{9z}p\int te^{-fz}dt-\frac{4}{9}p^2\int t^2e^{-fz}\right)=0 
\end{align}

Notice that 
\begin{align*}
\frac{4}{9}p^2\int t^2e^{-fz}=\frac{4}{27a_1}p^2\int(3a_1t^2+p)e^{-fz}-\frac{4}{27a_1}p^3I=-\frac{4}{27a_1}p^3I
\end{align*}

\begin{align*}
-\frac{10}{9z}p\int te^{-fz}dt=\frac{5}{9z}\int pte^{-fz}-\frac{5}{3z}\int(pt+q)e^{-fz}+\frac{5}{3z}qI=\\
\frac{5}{9z}\int pte^{-fz}-\frac{5}{3z}\int(pt+q+a_1t^3)e^{-fz}+\frac{5}{3z}\int a_1t^3e^{-fz}+\frac{5}{3z}qI=\\
\frac{5}{9z}\int t(3a_1t^2+p)e^{-fz}+\frac{5}{3z}I'+\frac{5}{3z}qI\\
=\frac{5}{9z^2}I+\frac{5}{3z}I'+\frac{5}{3z}qI
\end{align*}
thefore, collecting all the contributions together we find

\begin{align*}
I''+q^2I+2qI'-\frac{2}{3z}I'-\frac{2q}{3z}I-\frac{2}{3z^2}I+\frac{5}{9z^2}I+\frac{5}{3z}I'+\frac{5}{3z}qI+\frac{4}{27a_1}p^3I=0\\
I''+2qI'+\left(\frac{4p^3}{27a_1}+q^2\right)I+\frac{1}{z}I'+\frac{q}{z}I-\frac{1}{9z^2}I=0
\end{align*}

Case $(2)$: if $p=0$, then integrating by part we have
\begin{align*}
I(z)&=\int_{\mathcal{C}_1+\frac{a_2}{3a_1}}e^{-(a_1t^3+q)z}dt\\
&=\left[te^{-(a_1t^3+q)z}\right]_{\mathcal{C}_1+\frac{a_2}{3a_1}}+\int_{\mathcal{C}_1+\frac{a_2}{3a_1}}3a_1t^3ze^{-(a_1t^3+q)z}dt\\
&=3z\int_{\mathcal{C}_1+\frac{a_2}{3a_1}}(a_1t^3+q)e^{-(a_1t^3+q)z}dt-3qz\int_{\mathcal{C}_1+\frac{a_2}{3a_1}}e^{-(a_1t^3+q)z}dt\\
&=-3zI'(z)-3qzI(z)
\end{align*}
\end{proof}
%  \[I(z)=\int_{\mathcal{C}_{\alpha}}e^{-zf}dt\] is a solution of a second order ODE with regular/irregular singularity at zero/infinity. In particular, only two cases can occur: let $f(t)=a_1t^3+a_2t^2+a_3t+a_4$
%\begin{enumerate}
%\item if $a_2=0$ and $a_3\neq 0$, \begin{equation} I''+2a_4 I'+\left(a_4^2+\frac{4}{27}\frac{a_3^3}{a_1}\right)I+\frac{I'}{z}+\frac{a_4}{z}I-\frac{1}{9z^2}=0
%\end{equation}
%\item if $a_2=a_3=0$, $I$ solves
%\begin{equation}\label{eq:I,a,e}
%I''+aI'+(aa_4-a_4^2)I+\frac{4+9e}{3}\frac{I'}{z}+\frac{a+2a_4+9a_4e}{3}\frac{I}{z}+\frac{e}{z^2}I=0
%\end{equation}
%for every $a,e\in\C$; in particular if $I$ is a solution of
%\begin{equation}
%I'+\left(a_4+\frac{1}{3z}\right)I=0
%\end{equation}
%then it solves equation \eqref{eq:I,a,e} for every $a,e\in\C$.   
%\end{enumerate}  
%We notice that in the first case (of which Airy is an example), since $a_3\neq 0$ $f(t)$ has always two distinct critical points. Wherease, if $a_3=0$ there is a single critical point at $t=0$.

We would like to verify that for every cubic function $f$, the Borel transform of the exponential integral can be expressed by an hypergeometric function and hence deduce its resurgent properties in full generality. If $p\neq 0$, $I(z)$ is a solution of

\begin{equation}
I''+2qI'+\left(\frac{4p^3}{27a_1}+q^2\right)I+\frac{1}{z}I'+\frac{q}{z}I-\frac{1}{9z^2}I=0
\end{equation}

hence a formal solution as $z\to \infty$ is given (up to constants $U_1,U_2\in\C$) by 
\begin{align}
\tilde{I}_{+}(z)&\defeq U_1e^{-(q+\sqrt{\frac{4p^3}{27a_1}})z}z^{1/2}\tilde{w}_+(z)\\
\tilde{I}_-(z)&\defeq U_2 e^{-(q-\sqrt{\frac{4p^3}{27a_1}})z}z^{1/2}\tilde{w}_-(z)
\end{align}
where $\tilde{w}_{\pm}(z)\in\C[\![z^{-1}]\!]$ is the formal solution of
\begin{equation}\label{w_general}
\tilde{w}_{\pm}''\mp 2\sqrt{\frac{4p^3}{27a_1}}\tilde{w}_{\pm}'+\frac{5}{36}\frac{\tilde{w}_{\pm}}{z^2}=0
\end{equation}
with $\tilde{w}_{\pm}(z)=1+\sum_{k\geq 1}a_{\pm,k}z^{-k}$.

We can now compute the Borel transform of \eqref{w_general}: for $\tilde{w}_+(z)$

\begin{align*}
\zeta^2\hat{w}-2\sqrt{\frac{4p^3}{27a_1}}\zeta\hat{w}+\frac{5}{36}\int_0^{\zeta}(\zeta-\zeta')\hat{w}(\zeta')d\zeta'&=0 &\\
\zeta^2\hat{w}'+2\zeta\hat{w}+2\sqrt{\frac{4p^3}{27a_1}}\hat{w}+2\sqrt{\frac{4p^3}{27a_1}}\zeta\hat{w}'+\frac{5}{36}\int_0^\zeta\hat{w}(\zeta')&=0 &\\
\left(\zeta^2+2\sqrt{\frac{4p^3}{27a_1}}\zeta\right)\hat{w}''+4\left(\zeta+\sqrt{\frac{4p^3}{27a_1}}\right)\hat{w}'+\frac{77}{36}\hat{w}&=0 &\\
t(1-t)\hat{w}''+(2-4t)\hat{w}'-\frac{77}{36}\hat{w}&=0 &\zeta=-2t\sqrt{\frac{4p^3}{27a_1}}
\end{align*}

hence \[\hat{w}_+(\zeta)={}_2F_1\left(\frac{7}{6},\frac{11}{6};2;-\frac{3}{4p}\sqrt{\frac{3a_1}{p}}\zeta\right)\] and analougusly, \[\hat{w}_-(\zeta)={}_2F_1\left(\frac{7}{6},\frac{11}{6};2;\frac{3}{4p}\sqrt{\frac{3a_1}{p}}\zeta\right).\] Notice that $\hat{w}_{\pm}(\zeta)$ has a branch cut singularity respectively at $\zeta=\zeta_{\pm}\defeq\pm \sqrt{\frac{16p^3}{27a_1}}$, and thanks to the well known formulas for the analytic continuation of hypergeometric functions (see 15.2.3 DLMF), if we assume the branch cut is from $\zeta_{\pm}$ to $+\infty$

\begin{align*}
\hat{w}_-(\zeta+i0)-\hat{w}_-(\zeta-i0)&=\frac{2\pi i}{\Gamma(7/6)\Gamma(11/6)}\left(\frac{\zeta}{\zeta_+}-1\right)^{-1}\sum_{k\geq 0}\frac{(5/6)_k(1/6)_k}{\Gamma(k)k!}\left(1-\frac{\zeta}{\zeta_+}\right)^k & \zeta\in\left(\zeta_+,+\infty\right) & \\
&=-\frac{2\pi i}{\Gamma(7/6)\Gamma(11/6)} \sum_{k\geq 0}\frac{(5/6)_k(1/6)_k}{\Gamma(k)k!}\left(1-\frac{\zeta}{\zeta_+}\right)^{k-1} & \\
&=\mathbf{-i}\frac{1}{\Gamma(7/6)\Gamma(11/6)}\sum_{k\geq 0}\frac{\Gamma\left(\tfrac{7}{6}+k\right)\Gamma\left(\tfrac{11}{6}+k\right)}{\Gamma(2+k)\Gamma(k+1)}\left(1-\frac{\zeta}{\zeta_+}\right)^k & \\
&=\mathbf{-i}\,\, {}_2F_1\left(\frac{7}{6},\frac{11}{6};2;1-\frac{\zeta}{\zeta_+}\right) & \\
&=\mathbf{-i}\hat{w}_+(\zeta-\zeta_+)
\end{align*}

Similarly, if we assume that the branch cut is from $\zeta_{\pm}$ to $-\infty$ then

\begin{align*}
\hat{w}_+(\zeta+i0)-\hat{w}_+(\zeta-i0)&=-\frac{2\pi i}{\Gamma(7/6)\Gamma(11/6)}\left(\frac{\zeta}{\zeta_-}-1\right)^{-1}\sum_{k\geq 0}\frac{(5/6)_k(1/6)_k}{\Gamma(k)k!}\left(1-\frac{\zeta}{\zeta_-}\right)^k & \zeta\in\left(-\infty,\zeta_-\right) & \\
&=\frac{2\pi i}{\Gamma(7/6)\Gamma(11/6)} \sum_{k\geq 0}\frac{(5/6)_k(1/6)_k}{\Gamma(k)k!}\left(1-\frac{\zeta}{\zeta_-}\right)^{k-1} & \\
&=\mathbf{i}\frac{1}{\Gamma(7/6)\Gamma(11/6)}\sum_{k\geq 0}\frac{\Gamma\left(\tfrac{7}{6}+k\right)\Gamma\left(\tfrac{11}{6}+k\right)}{\Gamma(2+k)\Gamma(k+1)}\left(1-\frac{\zeta}{\zeta_-}\right)^k & \\
&=\mathbf{i}\,\, {}_2F_1\left(\frac{7}{6},\frac{11}{6};2;1-\frac{\zeta}{\zeta_-}\right) & \\
&=\mathbf{i}\hat{w}_-(\zeta-\zeta_-)
\end{align*}

therefore we see that the Stokes factors are given by $\pm i$ (as for Airy).

\textbf{I think it will be nice to add the geometric interpretation of Maxim in term of Lefschetz thimbles}

Let us first compute the Borel transform of \eqref{eq:I for p not 0} (indeed as in the proof of Theorem \ref{thm:maxim} we know that \eqref{eq:I for p not 0} admits a formal solution which is Gevrey-$1$)

\begin{align*}
\zeta^2\hat{I}-a\zeta\hat{I}+b\hat{I}-\int_0^{\zeta}\zeta'\hat{I}(\zeta')+d\int_0^{\zeta}\hat{I}(\zeta')-\frac{1}{9}\int_0^{\zeta}(\zeta-\zeta')\hat{I}(\zeta')=0\\
2\zeta\hat{I}+\zeta^2\hat{I}'-a\hat{I}-a\zeta\hat{I}'+b\hat{I}'-\zeta\hat{I}+d\hat{I}-\frac{1}{9}\int\hat{I}(\zeta')=0\\
(\zeta^2-a\zeta+b)\hat{I}''+(3\zeta-2a+d)\hat{I}'+\frac{8}{9}\hat{I}=0
\end{align*}

Now we denote by $\lambda_1,\lambda_2$ the distinguished (we assume that $p\neq 0$) roots of $\zeta^2-a\zeta+b$, then

\begin{align}
& (\zeta-\lambda_1)(\zeta-\lambda_2)\hat{I}''+(3\zeta-2a+d)\hat{I}'+\frac{8}{9}\hat{I}=0 & \\
&(t+\lambda_2-\lambda_1)t\hat{I}''+(3t+3\lambda_2-2a+d)\hat{I}'+\frac{8}{9}=0 & t=\zeta-\lambda_2 \\
& \label{eq:I hyper} s(1-s)\hat{I}''-\left(3s+\frac{3\lambda_2-2a+d}{\lambda_1-\lambda_2}\right)\hat{I}'-\frac{8}{9}\hat{I}=0 & t=(\lambda_1-\lambda_2)s
\end{align}

where \eqref{eq:I hyper} is an hypergeometric equation\footnote{Notice that $\lambda_{1,2}=q\pm \frac{2i}{3}p\sqrt{\frac{p}{3a_1}}$, $a=2q$ and $d=q$. Hence \[\frac{2a-d-3\lambda_2}{\lambda_1-\lambda_2}=\frac{4q-q-3q-2ip\sqrt{\frac{p}{3a_1}}}{-\frac{4i}{3}p\sqrt{\frac{p}{3a_1}}}=\frac{3}{2}\]} and a solution is given by

\begin{equation}
\hat{I}_{\lambda_1}(\zeta;U_1,U_2)=U_1\,{}_2F_1\left(\frac{2}{3},\frac{4}{3};\frac{3}{2};\frac{\zeta-\lambda_2}{\lambda_1-\lambda_2}\right)+U_2\left(\frac{\zeta-\lambda_2}{\lambda_1-\lambda_2}\right)^{-1/2}{}_2F_1\left(\frac{1}{6},\frac{5}{6};\frac{1}{2};\frac{\zeta-\lambda_2}{\lambda_1-\lambda_2}\right)
\end{equation}

which has a branch cut at $\zeta=\lambda_1$, where $U_1,U_2$ are constants. Of course, reversing the role of $\lambda_1$ and $\lambda_2$ we find

\begin{equation}
\hat{I}_{\lambda_2}(\zeta,;U_1,U_2)=U_1{}_2F_1\left(\frac{2}{3},\frac{4}{3};\frac{3}{2};\frac{\zeta-\lambda_1}{\lambda_2-\lambda_1}\right)+U_2\left(\frac{\zeta-\lambda_1}{\lambda_2-\lambda_1}\right)^{-1/2}{}_2F_1\left(\frac{1}{6},\frac{5}{6};\frac{1}{2};\frac{\zeta-\lambda_1}{\lambda_2-\lambda_1}\right)
\end{equation}

is the Borel transform of $\tilde{I}_{\lambda_2}(z)$ and it has a branch cut singularity at $\zeta=\lambda_2$. It is remarkable that  the dependece on the function $f$ is only on the location of the singularities, but it is always an hypergeometric function with the same parameters. In addition, we can compute the Stokes constants thanks to the well known formula for analyitic continuation of hypergeometric (see 15.2.3 in DLMF)

\begin{align*}
\hat{I}_{\lambda_1}(\zeta+i0;U_1,0)-\hat{I}_{\lambda_1}(\zeta-i0;U_1,0)&=-U_1\frac{2\pi i}{\Gamma\left(\frac{2}{3}\right)\Gamma\left(\frac{4}{3}\right)}\left(\frac{\lambda_1-\zeta}{\lambda_2-\lambda_1}\right)^{-1/2}{}_2F_1\left(\frac{1}{6},\frac{5}{6};\frac{1}{2};\frac{\zeta-\lambda_1}{\lambda_2-\lambda_1}\right)\\
&=\mathbf{-i}U_1\frac{2\pi i}{\Gamma\left(\frac{2}{3}\right)\Gamma\left(\frac{4}{3}\right)}\left(\frac{\zeta-\lambda_1}{\lambda_2-\lambda_1}\right)^{-1/2}{}_2F_1\left(\frac{1}{6},\frac{5}{6};\frac{1}{2};\frac{\zeta-\lambda_1}{\lambda_2-\lambda_1}\right)\\
&=\mathbf{-i} \hat{I}_{\lambda_2}\left(\zeta;0,\frac{U_1}{\Gamma(2/3)\Gamma(4/3)}\right)
\end{align*}

\end{document}