### 2022-02-23 modified bessel.tex
update modified Bessel

\documentclass{article}

\usepackage{url}
\usepackage[hmargin=1.5in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{xcolor}
% convenience aliases
\newcommand{\maps}{\colon}

% symbology
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\laplace}{\mathcal{L}}
\include{command}

\title{Resurgence of modified Bessel functions of second kind}
\author{Veronica Fantini}

\begin{document}
\maketitle

\section{Modified Bessel function of second kind}

The modified Bessel function of the second kind $K_\mu(z)$ is a solution of the equation
\begin{equation}\label{Bessel_nu}
w''+\frac{w'}{z}-w-\frac{\mu^2}{z^2}=0
\end{equation}
such that $K_{\mu}(z)\sim\sqrt{\pi/(2z)}e^{-z}$ as $z\to\infty$ in $|\arg z|<\frac{3\pi}{2}$\footnote{A system of solution of Bessel equation is given by $I_\mu(z)$ and $K_\mu(z)$. In particular, their asymptotic behaviour as $z\to\infty$ is given by \begin{align}
\tilde{I}_\mu(z)&=\frac{1}{\sqrt{2\pi}}e^zz^{-1/2}\sum_{k\geq 0}\frac{\left(\frac{1}{2}-\mu\right)_k\left(\frac{1}{2}+\mu\right)_k}{2^kk!}z^{-k}\\
\tilde{K}_{\mu}(z)&=\sqrt{\frac{\pi}{2}}e^{-z}z^{-1/2}\sum_{k\geq 0}\frac{\left(\frac{1}{2}-\mu\right)_k\left(\frac{1}{2}+\mu\right)_k}{(-2)^kk!}z^{-k}
\end{align}}. It has a branch point at $z=0$ for every $\mu\in\C$ and  the principal branch is analytic in $\C\setminus(-\infty,0]$. 
\subsection{Differential equation}
From the general theory of ODE, the formal integral solution of \eqref{Bessel_nu} is two parameters family 
\begin{equation}
\tilde{w}(z)=U_1e^{-z}z^{-1/2}\tilde{w}_{\mu,+}(z)+U_2e^{z}z^{-1/2}\tilde{w}_{\mu,-}(z)
\end{equation} 
where $\tilde{w}_{\mu,\pm}=\sum_{j\geq 0}a_{\pm,j}z^{-j}\in\C[\![z^{-1}]\!]$ are unique formal solutions of
\begin{align*}
\tilde{w}_{\mu,+}''-2\tilde{w}_{\nu,+}'+\frac{\tilde{w}_{\mu,+}}{4z^2}-\frac{\mu^2}{z^2}\tilde{w}_{\mu,+}=0\\
\tilde{w}_{\mu,-}''+2\tilde{w}_{\mu,-}'+\frac{\tilde{w}_{\mu,-}}{4z^2}-\frac{\mu^2}{z^2}\tilde{w}_{\mu,-}=0
\end{align*} 
In particular, $\tilde{K}_{\mu}(z)=\sqrt{\frac{\pi}{2}}e^{-z}z^{-1/2}\tilde{w}_{\mu,+}(z)$ and $\tilde{I}_\mu(z)=\frac{1}{\sqrt{2\pi}}e^zz^{-1/2}\tilde{w}_{\mu,-}(z)$ (once we choose $a_{\pm,0}=1$). We now compute the Borel transform of $\tilde{w}_+(z)$\footnote{We do not consider constant term of $\tilde{w}_{\mu,\pm}$, i.e. $\mathcal{B}:\C[\![z^{-1}]\!]\to \C[\zeta]$.} it is a solution of 
\begin{align*}
&\zeta^2\hat{w}_{\mu,+}+2t\hat{w}_{\mu,+}+\left(\frac{1}{4}-\nu^2\right)\int_0^{\zeta}(\zeta-s)\hat{w}_{\nu,+}(s)ds=0 &\\
&\zeta^2\hat{w}_{\mu,+}''+2\zeta\hat{w}_+''+4\zeta\hat{w}_{\mu,+}'+\left(\frac{9}{4}-\mu^2\right)\hat{w}_{\mu,+}=0 & \\
&t(1-t)\hat{w}_{\mu,+}''+(2-4t)\hat{w}_{\mu,+}'-\left(\frac{9}{4}-\mu^2\right)\hat{w}_{mu,+}=0 & t=-\frac{\zeta}{2}
\end{align*} 
therefore $\hat{w}_{\mu,+}(\zeta)$ is an hypergeometric function
\begin{equation}
\hat{w}_{\mu,+}(\zeta)={}_2F_1\left(\frac{3}{2}-\mu,\frac{3}{2}+\mu;2;-\frac{\zeta}{2}\right)
\end{equation}
and it has a branch point singularities at $\zeta=-2$. By the same reasoning, 
\begin{equation}
\hat{w}_{\mu,-}(\zeta)={}_2F_1\left(\frac{3}{2}-\mu,\frac{3}{2}+\mu;2;\frac{\zeta}{2}\right)
\end{equation} 
and it has branch point at $\zeta=2$. Thanks to explicit formula for the analytic continuation of hypergeomtric functions (see \cite{dlmf} 15.2.3)
\begin{align*}
\textcolor{red}{\frac{\pi\sqrt{\pi}}{\cos(\mu \pi)}}\left(\hat{w}_{\mu,+}(\zeta+i0)-\hat{w}_{\mu,+}(\zeta-i0)\right)&=\textcolor{red}{\frac{\pi\sqrt{\pi}}{\cos(\mu \pi)}}\frac{2\pi i}{\Gamma(\tfrac{3}{2}-\mu)\Gamma(\tfrac{3}{2}+\mu)}\left(-\frac{\zeta}{2}-1\right)^{-1}{}_2F_1\left(\frac{1}{2}+\mu,\frac{1}{2}-\mu;0;1+\frac{\zeta}{2}\right) & \zeta>-2 \\
&=-\textcolor{red}{\frac{\pi\sqrt{\pi}}{\cos(\mu \pi)}}\frac{2\pi i}{\Gamma(\tfrac{3}{2}-\mu)\Gamma(\tfrac{3}{2}+\mu)}\sum_{k\geq 0}\frac{\left(\tfrac{1}{2}-\mu\right)_k\left(\tfrac{1}{2}+\mu\right)_k}{\Gamma(k)k!}\left(1+\frac{\zeta}{2}\right)^{k-1} & \\
&=-\textcolor{red}{\frac{\pi\sqrt{\pi}}{\cos(\mu \pi)}}\frac{2\pi i}{\Gamma(\tfrac{3}{2}-\mu)\Gamma(\tfrac{3}{2}+\mu)}\sum_{k\geq 1}\frac{\left(\tfrac{1}{2}-\mu\right)_k\left(\tfrac{1}{2}+\mu\right)_k}{\Gamma(k)k!}\left(1+\frac{\zeta}{2}\right)^{k-1} & \\
&=-\textcolor{red}{\frac{\pi\sqrt{\pi}}{\cos(\mu \pi)}}\frac{2\pi i}{\Gamma(\tfrac{3}{2}-\mu)\Gamma(\tfrac{3}{2}+\mu)}\frac{1}{\Gamma\left(\frac{1}{2}-\mu\right)\Gamma\left(\frac{1}{2}+\mu\right)}\cdot\\
& \qquad\cdot \sum_{k\geq 1}\frac{\Gamma\left(\tfrac{1}{2}-\mu+k\right)\Gamma\left(\tfrac{1}{2}+\mu+k\right)}{\Gamma(k)k!}\left(1+\frac{\zeta}{2}\right)^{k-1} & \\
&=-\textcolor{red}{\frac{\pi\sqrt{\pi}}{\cos(\mu \pi)}}\frac{2\pi i}{\Gamma(\tfrac{3}{2}-\mu)\Gamma(\tfrac{3}{2}+\mu)}\frac{1}{\Gamma\left(\frac{1}{2}-\mu\right)\Gamma\left(\frac{1}{2}+\mu\right)}\cdot \\
&\qquad\cdot \sum_{k\geq 0}\frac{\Gamma\left(\tfrac{3}{2}-\mu+k\right)\Gamma\left(\tfrac{3}{2}+\mu+k\right)}{\Gamma(k+1)(k+1)!}\left(1+\frac{\zeta}{2}\right)^{k} & \\
&=-\textcolor{red}{\frac{\pi\sqrt{\pi}}{\cos(\mu \pi)}}\frac{2\pi i}{\Gamma\left(\frac{1}{2}-\mu\right)\Gamma\left(\frac{1}{2}+\mu\right)}\,\, {}_2F_1\left(\frac{3}{2}-\mu,\frac{3}{2}+\mu;2;1+\frac{\zeta}{2}\right) & \\
&=-2i\textcolor{red}{\sqrt{\pi}}\hat{w}_{\mu,-}(\zeta+2)
\end{align*} 
and for $\hat{w}_{\mu,-}(\zeta)$
\begin{align*}
\hat{w}_{\mu,-}(\zeta+i0)-\hat{w}_{\mu,-}(\zeta-i0)&=-\frac{2\pi i}{\Gamma(\tfrac{3}{2}-\mu)\Gamma(\tfrac{3}{2}+\mu)}\left(\frac{\zeta}{2}-1\right)^{-1}{}_2F_1\left(\frac{1}{2}-\mu,\frac{1}{2}+\mu;0;1-\frac{\zeta}{2}\right) & \zeta<2 \\
&=\frac{2\pi i}{\Gamma(\tfrac{3}{2}-\mu)\Gamma(\tfrac{3}{2}+\mu)}\sum_{k\geq 0}\frac{\left(\tfrac{1}{2}-\mu\right)_k\left(\tfrac{1}{2}+\mu\right)_k}{\Gamma(k)k!}\left(1-\frac{\zeta}{2}\right)^{k-1} & \\
&=2i\cos(\nu \pi)\,\, {}_2F_1\left(\frac{3}{2}-\mu,\frac{3}{2}+\mu;2;1-\frac{\zeta}{2}\right) & \\
&=2i\cos(\mu \pi)\hat{w}_{\mu,+}(\zeta-2)
\end{align*}
In addition, the previous relations computes the Stokes constants which are funtions of $\nu$ and are given by $\pm 2i\cos(\mu\pi)$.

\subsection{Exponential integral}

Let $X=\C^*$, $f(x)=x+\frac{1}{x}$ and for every $\mu\in [0,+\infty)$ let $\nu=x^{\mu-1}+\frac{1}{x^{\mu+1}}dx$, then

\begin{equation}
I(z;m)\defeq\int_{0}^{\infty}e^{-zf}\nu
\end{equation}

In particulr, on the universal cover $\pi\colon\tilde{C}\to \C^*$ setting $x=e^u$ 
\begin{equation}
I(\frac{z}{2};\mu)=2\int_{-\infty}^{\infty}e^{-z\cosh(u)}\cosh(\mu u)du=4K_\mu(z)\,\quad |\arg(z)|<\pi/2
\end{equation} 
where $K_\mu(z)$ is the second kind modified Bessel function with parameter $\mu$. It is worth mentioning that $I(z;\mu)$ differs from $I(z;0)$ only in $\pi^*(\nu)$ while $\pi^*(f)$ stays the same for every $\mu\in [0,\infty)$. Hence we can adapt part of the argument used in Bessel example (see ), and verify the $3/2$-derivative formula: let $\zeta=\cosh(u)$

\begin{align*}
\int_{\mathcal{C}_0(\zeta)}\pi^*(\nu)&=\int_{\mathcal{C}_0(\zeta)}\cosh(\mu u)du\\
&=\frac{1}{\mu}\Big[\sinh(\mu u)\Big]_{\mathrm{start}\mathcal{C}_0(\zeta)}^{\mathrm{end}\mathcal{C}_0(\zeta)}\\
&=\frac{1}{\mu}\left(\sinh\left(\mu\,\mathrm{acosh}\left(\zeta\right)\right)-\sinh\left(-\mu\,\mathrm{acosh}\left(\zeta\right)\right)\right)\\
&=\frac{2}{\mu}\sinh\left(\mu\,\mathrm{acosh}\left(\zeta\right)\right)
\end{align*}

The we set $\xi=\frac{1}{2}\left(\zeta-1\right)$, thanks to identity 15.4.16 \textbf{DLMF}

\begin{align*}
\sinh(\tau) {}_2F_1\left(\frac{1}{2}-\mu,\frac{1}{2}+\mu;\frac{3}{2};-\sinh^2(\tau)\right)&=\frac{1}{2\mu}\sinh(2\mu\tau) & \\
\xi^{1/2} {}_2F_1\left(\frac{1}{2}-\mu,\frac{1}{2}+\mu;\frac{3}{2};-\xi\right)&=\frac{1}{2\mu}\sinh(2\mu\tau)  & \sinh^2(\tau)=\xi \\
\xi^{1/2} {}_2F_1\left(\frac{1}{2}-\mu,\frac{1}{2}+\mu;\frac{3}{2};-\xi\right)&=\frac{1}{2\mu}\sinh(\mu\,\,\mathrm{acosh}\left(\zeta\right))  & \cosh(2\tau)=\zeta \\
&=\frac{1}{4}\int_{\mathcal{C}_0(\zeta)}\pi^*(\nu) &\\
\end{align*}

Thus we take $3/2$-derivative based at $\zeta=1$

\begin{align*}
\partial_{\zeta}^{3/2}\left(\int_{\mathcal{C}_0(\zeta)}\pi^*(\nu)\right)&=\partial_\zeta^2\left(\frac{1}{\Gamma\left(\frac{1}{2}\right)}\int_1^\zeta(\zeta-\zeta')^{-1/2}\left(\int_{\mathcal{C}_0(\zeta')}\pi^*(\nu)\right)d\zeta'\right)\\
&=4\partial_\zeta^2\left(\frac{1}{\Gamma\left(\frac{1}{2}\right)}\int_0^\xi\frac{1}{2}(\xi-\xi')^{-1/2}\xi^{1/2} {}_2F_1\left(\frac{1}{2}-\mu,\frac{1}{2}+\mu;\frac{3}{2};-\xi'\right) 4\,\,d\xi'\right)\\
&=8\partial_\zeta^2\left(\Gamma\left(\frac{3}{2}\right)\xi\,\,{}_2F_1\left(\frac{1}{2}-\mu,\frac{1}{2}+\mu;2;-\xi\right)\right) \\
&=4\sqrt{\pi}\frac{1}{4}\partial_\xi^2\left(\xi\,\,{}_2F_1\left(\frac{1}{2}-\mu,\frac{1}{2}+\mu;2;-\xi\right)\right)\\
&=-\sqrt{\pi}\partial_{\xi}\,\, {}_2F_1\left(\frac{1}{2}-\mu,\frac{1}{2}+\mu;1;-\xi\right)\\
&=\sqrt{\pi}\Gamma\left(\frac{1}{2}-\mu\right)\Gamma\left(\frac{1}{2}+\mu\right){}_2F_1\left(\frac{3}{2}-\mu,\frac{3}{2}+\mu;2;-\xi\right)\\
&=\frac{\pi\sqrt{\pi}}{\cos(\mu \pi)}{}_2F_1\left(\frac{3}{2}-\mu,\frac{3}{2}+\mu;2;\frac{1}{2}-\frac{\zeta}{2}\right)\\
&=\frac{\pi\sqrt{\pi}}{\cos(\mu \pi)}\hat{w}_{\mu,+}(\zeta-1)
\end{align*}

\color{orange}
As showed by Aaron, if $T_n(u)$ and $U_n(u)$ denote the Chebyschev polynomials \footnote{$T_n(\cos(t))=\cos(nt)$ and $U_n(\cos(t))\sin(t)=\sin((n+1)t)$.}

\begin{align*}
K_{\nu}(z)&=\frac{1}{2 i\sin(\nu \pi)}\int_{\mathcal{C}_{\alpha}}e^{z\cosh(t)}\sinh(\nu t) dt & u=\cosh(\nu t)\\
&=-\frac{1}{2i\nu\sin(\nu\pi)}\int_{\mathcal{C}_{\alpha}}e^{zT_{\frac{1}{\nu}}(u)}du & \zeta=T_{\frac{1}{\nu}}(u)\\
&=\frac{1}{2i\sin(\nu\pi)}\int_{\mathcal{H}_{\alpha}}e^{-\zeta z}\frac{d\zeta}{U_{\frac{1}{\nu}-1}(u)}\\
&=-\frac{1}{2i\sin(\nu\pi)}\int_{\mathcal{H}_{\alpha}}e^{-\zeta z}{}_2F_1\left(\frac{1-\nu}{2},\frac{1+\nu}{2};\frac{3}{2};1-\zeta^2\right)d\zeta
\end{align*}
where $\mathcal{C}_{\alpha}$ \textbf{has to be checked, but I guess } $\alpha=1$

Indentity 15.10.17 from \cite{dlmf} splits the integrand above into
\begin{align*}
{}_2F_1\left(\frac{1-\nu}{2},\frac{1+\nu}{2};\frac{3}{2};1-\zeta^2\right)&=C_1 \, {}_2F_1\left(\frac{1-\nu}{2},\frac{1+\nu}{2};\frac{1}{2};\zeta^2\right)+C_2\zeta \,{}_2F_1\left(1-\frac{\nu}{2},1+\frac{\nu}{2};\frac{3}{2};\zeta^2\right)\\
&=\tilde{C}_1\left[{}_2F_1\left({1-\nu},{1+\nu};\frac{3}{2};\frac{1}{2}-\frac{\zeta}{2}\right)+{}_2F_1\left({1-\nu},{1+\nu};\frac{3}{2};\frac{1}{2}+\frac{\zeta}{2}\right)\right]+\\
&\qquad+\tilde{C}_2\left[{}_2F_1\left({1-\nu},{1+\nu};\frac{3}{2};\frac{1}{2}-\frac{\zeta}{2}\right)+{}_2F_1\left({1-\nu},{1+\nu};\frac{3}{2};\frac{1}{2}+\frac{\zeta}{2}\right)\right]
\end{align*}
therefore, collecting the contrubutions together we have 
\begin{multline}
K_{\nu}(z)=\frac{i}{4\sin(\nu\pi)}\int_{\mathcal{H}_{\alpha}}e^{-\zeta z}\left[{}_2F_1\left({1-\nu},{1+\nu};\frac{3}{2};\frac{1}{2}-\frac{\zeta}{2}\right)+{}_2F_1\left({1-\nu},{1+\nu};\frac{3}{2};\frac{1}{2}+\frac{\zeta}{2}\right)\right] d\zeta
\end{multline}

Since ${}_2F_1\left({1-\nu},{1+\nu};\frac{3}{2};\frac{1}{2}+\frac{\zeta}{2}\right) $ is singular at $\zeta=1$, the inverse Laplace transform of $K_\nu(z)$ is 
\[
\hat{K}_{\nu}(\zeta)=\frac{i}{4\sin(\nu\pi)}{}_2F_1\left({1-\nu},{1+\nu};\frac{3}{2};\frac{1}{2}-\frac{\zeta}{2}\right)
\] 
\color{black}

\bibliographystyle{utphys}
\bibliography{airy-resurgence}
\end{document}