### 2022-11-28 hypergeometric.tex

\DeclareSymbolFont{AMSb}{U}{msb}{m}{n}
\documentclass[11pt,a4paper,twoside,leqno,noamsfonts]{amsart}
           \usepackage{setspace}
\linespread{1.34}           
           %\onehalfspacing
\usepackage[english]{babel}
\usepackage[dvipsnames]{xcolor}
\definecolor{britishracinggreen}{rgb}{0.0, 0.26, 0.15}
\definecolor{cobalt}{rgb}{0.0, 0.28, 0.67}
\usepackage[utopia]{mathdesign}
    \DeclareSymbolFont{usualmathcal}{OMS}{cmsy}{m}{n}
    \DeclareSymbolFontAlphabet{\mathcal}{usualmathcal}
\usepackage[a4paper,top=4cm,bottom=3cm,left=3.5cm,
           right=3.5cm,bindingoffset=5mm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{braket,caption,comment,mathtools,stmaryrd}
\usepackage{multirow,booktabs,microtype}
\usepackage{latexsym}
\usepackage{todonotes}
\usepackage{fancyhdr}
%\renewcommand{\sectionmark}[1]{\markboth{\thesection\ #1}{}}
\pagestyle{fancy}
% Clear the header and footer
\fancyhead{}
\fancyfoot{}
% Set the right side of the footer to be the page number
\fancyfoot[R]{\thepage}
\addtolength{\headheight}{\baselineskip}
%\fancyhead[RE]{\rightmark}
%\fancyhead[RE]{}
\usepackage{soul} % per testo barrato
\usepackage[colorlinks,bookmarks]{hyperref} %
\hypersetup{colorlinks,%
            citecolor=britishracinggreen,%
            filecolor=black,%
            linkcolor=cobalt,%
            urlcolor=black}
\setcounter{tocdepth}{2}
%\setcounter{section}{-1}
\numberwithin{equation}{section}

% Veronica's custom commands
\input{command}

% symbology
\newcommand{\blankbox}{{\fboxsep 0pt \colorbox{lightgray}{\phantom{$h$}}}}
\newcommand{\maps}{\colon}
\newcommand{\van}{\mathfrak{m}}
\newcommand{\laplace}{\mathcal{L}}
\newcommand{\borel}{\mathcal{B}}
\newcommand{\laplacepde}{\mathcal{D}}
\DeclareMathOperator{\Ai}{Ai}

\DeclareRobustCommand{\subtitle}[1]{\\#1}

\title{Resurgence of the Airy function \\ and other exponential integrals}
\author{Veronica Fantini and Aaron Fenyes}

\begin{document}
\maketitle
\section{Hypergeometric functions as Borel transform of second order ODE (\textit{series normales de \emph{I}er ordre})}

Let us consider the following linear second order ODE

\begin{equation}\label{eq}
\big[P(\tfrac{\partial}{\partial z})+\tfrac{1}{z}Q(\tfrac{\partial}{\partial z})+\tfrac{1}{z}R(\tfrac{1}{z})\big]f(z)=0
\end{equation}
with $\deg P=2$, $\deg Q=1$ and $R=O(\tfrac{1}{z})$. We denote by $\alpha_1,\alpha_2$ the roots of $P(-\lambda)$ and we assume they are distinct. Furthermore we assume $\tau_j:=\frac{Q(\alpha_j)}{P'(\alpha_j)}\neq 0$. The latter assumption guarantees the formal solution $\tilde{f}$ being slight, while the former assumption implies there will be two independent solutions.

Under the previous assumptions we prove that the Borel transformed solution $\hat{f}(\zeta_j)$ is a Gauss hypergeometric function, $\zeta_j=\zeta-\alpha_j$.

\begin{prop}
Let $P(\lambda)=\lambda^2+a_1\lambda+a_0$, $Q(\lambda)=b_1\lambda+b_0$ and $R(\tfrac{1}{z})=\tfrac{c_1}{z}$ satisfying the previous assumptions. Then 
\begin{align}
\hat{f}(\zeta_1)&=\left(1-\frac{\zeta_1}{\alpha_2-\alpha_1}\right)^{c-1} \, {}_2F_1\left(a,b;c;1-\frac{\zeta_1}{\alpha_2-\alpha_1} \right)\\
\hat{f}(\zeta_2)&=\left(1-\frac{\zeta_2}{\alpha_1-\alpha_2}\right)^{c-1} \, {}_2F_1\left(a,b;c;1-\frac{\zeta_2}{\alpha_1-\alpha_2} \right)
\end{align}
where the coefficients $a,b,c$ depend on the parameter of $P,Q,R$. 
\end{prop}

\begin{proof}
We start by taking the Borel transform of \eqref{eq}:
\begin{align}\label{eq:B}
(\zeta^2-a_1\zeta+a_0)\hat{f}(\zeta)+\int_0^\zeta b_1(-\zeta')\hat{f}(\zeta')d\zeta'+b_0\int_0^\zeta\hat{f}(\zeta')d\zeta'+c_1\int_0^\zeta(\zeta-\zeta')\hat{f}(\zeta')d\zeta'=0
\end{align} 
then we differentiate twice in order to have a differential equation which can be easier recognized as a hypergeometric equation. Since $\tilde{F}$ is slight and locally integrable at $0$ by assumption, Proposition 1 {\tt Resurgent Airy doc by Aaron} tells we are not loosing information taking derivatives, and that $\hat{f}(\zeta)$ is a solution of \eqref{eq:B} if and only if it is a solution of \eqref{eq:B_deriv}
\begin{align}
\label{eq:B_deriv}
\big[(\zeta^2-a_1\zeta+a_0)\partial_\zeta^2+(4\zeta-b_1\zeta-2a_1+b_0)\partial_\zeta+(c_1+2-b_1)\big]\hat{f}(\zeta)=0
\end{align}
We introduce some notation to simplify the computations, we denote by $\beta_1=4-b_1$, $\beta_0=b_0-2a_1$, $\gamma=c_1+2-b_1$ so \eqref{eq:B_deriv} turns into
\begin{align*}
\big[(\zeta-\alpha_1)(\zeta-\alpha_2)\partial_\zeta^2+(\beta_1\zeta+\beta_0)\partial_\zeta+\gamma\big]\hat{f}(\zeta)=0
\end{align*}
We consider the following change of coordinates $\zeta=\alpha_2-(\alpha_2-\alpha_1)\xi$\footnote{$\partial_\zeta=(\alpha_1-\alpha_2)^{-1}\partial_\xi$ and $\partial_\zeta^2=(\alpha_1-\alpha_2)^{-2}\partial_\xi^2$}
\begin{align*}
&\big[(\alpha_2-(\alpha_2-\alpha_1)\xi-\alpha_1)(\alpha_2-(\alpha_2-\alpha_1)\xi-\alpha_2)(\alpha_1-\alpha_2)^{-2}\partial_\xi^2+(\beta_1(\alpha_2-(\alpha_2-\alpha_1)\xi)+\beta_0)(\alpha_1-\alpha_2)^{-1}\partial_\xi+\gamma\big]\hat{f}(\xi)=0\\
&\big[(\alpha_2-\alpha_1)(1-\xi)(\alpha_1-\alpha_2)\xi(\alpha_1-\alpha_2)^{-2}\partial_\xi^2+(\beta_1\alpha_2-\beta_1(\alpha_2-\alpha_1)\xi+\beta_0)(\alpha_1-\alpha_2)^{-1}\partial_\xi+\gamma\big]\hat{f}(\xi)=0\\
&\big[-(1-\xi)\xi\partial_\xi^2+((\beta_1\alpha_2+\beta_0)(\alpha_1-\alpha_2)^{-1}+\beta_1\xi)\partial_\xi+\gamma\big]\hat{f}(\xi)=0\\
&\big[(1-\xi)\xi\partial_\xi^2-((\beta_1\alpha_2+\beta_0)(\alpha_1-\alpha_2)^{-1}+\beta_1\xi)\partial_\xi-\gamma\big]\hat{f}(\xi)=0
\end{align*}

The latter equation is an hypergeometric equation of parameters 
\begin{align*}
&C=(\beta_1\alpha_2+\beta_0)(\alpha_2-\alpha_1)^{-1}\\
&A+B+1=\beta_1=4-b_1 \Rightarrow A+B=3-b_1\\
&AB=\gamma=c_1+2-b_1
\end{align*}
and a solution is given by 
\begin{align*}
%\hat{f}(\xi)=(1-\xi)^{C-A-B} {}_2F_1(C-A,C-B;C;\xi)\\
\hat{f}(\xi)&=\xi^{1-C} {}_2F_1(A-C+1,B-C+1;2-C;\xi)\\
&=\left(\frac{\alpha_2-\zeta}{\alpha_2-\alpha_1}\right)^{1-C} {}_2F_1\left(A-C+1,B-C+1;2-C;\frac{\alpha_2-\zeta}{\alpha_2-\alpha_1}\right)\\
&=\left(1-\frac{\zeta_1}{\alpha_2-\alpha_1}\right)^{1-C} {}_2F_1\left(A-C+1,B-C+1;2-C;1-\frac{\zeta_1}{\alpha_2-\alpha_1}\right)
\end{align*}

\end{proof}

\begin{prop}
We verify that 
\begin{equation}
\hat{f}(\zeta_1)=\left(1-\frac{\zeta_1}{\alpha_2-\alpha_1}\right)^{1-C} \, {}_2F_1\left(A-C+1,B-C+1;2-C;1-\frac{\zeta_1}{\alpha_2-\alpha_1} \right)
\end{equation}
\begin{equation}
\hat{f}(\zeta_2)=\left(1-\frac{\zeta_2}{\alpha_1-\alpha_2}\right)^{1-C} \, {}_2F_1\left(A-C+1,B-C+1;2-C;1-\frac{\zeta_2}{\alpha_1-\alpha_2} \right)
\end{equation}
are indeed resurgent.  
\end{prop}

\begin{proof}
\color{Tan}
Recall that the analytic continuation of Gauss hypergeometric functions is given in terms of other hypergeometric fucntions (see DLMF 15.2.3):
\begin{align*}
\hat{f}(\zeta_1+i0)-\hat{f}(\zeta_1-i0)&\propto\left(\frac{\zeta_1}{\alpha_1-\alpha_2}\right)^{C-A-B}\left(1-\frac{\zeta_1}{\alpha_2-\alpha_1}\right)^{1-C} {}_2F_1\left(1-A,1-B;C+1-A-B;\frac{\zeta_1}{\alpha_2-\alpha_1}\right)\\
&=(-1)^{C-A-B}\left(1-\frac{\zeta_2}{\alpha_1-\alpha_2}\right)^{C-A-B}\left(\frac{\zeta_2}{\alpha_1-\alpha_2}\right)^{1-C}\cdot\\
&\qquad\qquad \cdot {}_2F_1\left(1-A,1-B;C+1-A-B;1-\frac{\zeta_2}{\alpha_1-\alpha_2}\right)\\
\end{align*}
\color{black}
\end{proof}

\section{Hypergeometric functions as Borel transform of third order ODE}

Let us consider the following linear third order ODE

\begin{equation}\label{eq3}
\big[P(\tfrac{\partial}{\partial z})+\tfrac{1}{z}Q(\tfrac{\partial}{\partial z})+\tfrac{1}{z}R(\tfrac{1}{z})\big]f(z)=0
\end{equation}
with $\deg P=3$, $\deg Q=2$ and $R=O(\tfrac{1}{z})$. We denote by $\alpha_1,\alpha_2,\alpha_3$ the roots of $P(-\lambda)$ and we assume they are distinct. There are three possible scenarios illustrated in figure \ref{fig3} below:
\begin{itemize}
\item $|\alpha_1-\alpha_2|<|\alpha_2-\alpha_3|<|\alpha_1-\alpha_3|$
\item $|\alpha_2-\alpha_3|=|\alpha_1-\alpha_3|<|\alpha_1-\alpha_2|$
\item $|\alpha_1-\alpha_2|=|\alpha_2-\alpha_3|=|\alpha_1-\alpha_3|$
\end{itemize}

\begin{figure}[h]\label{fig3}
\begin{tikzpicture}
\draw [red](-1,1)--(-3,1);
\draw [green](-1,1)--(-1.25,3);
\draw [blue](-3,1)--(-1.25,3);
\node[font=\tiny, below] at (-3,1) {$\alpha_1$};
\node[font=\tiny, below] at (-1,1) {$\alpha_2$};
\node[font=\tiny, above] at (-1.25,3) {$\alpha_3$};
\draw [green](0.25,1)--(1.5,2);
\draw [red](0.25,1)--(2.75,1);
\draw [green](1.5,2)--(2.75,1);
\node[font=\tiny, below] at (0.25,1) {$\alpha_1$};
\node[font=\tiny, below] at (2.75,1) {$\alpha_2$};
\node[font=\tiny, above] at (1.5,2) {$\alpha_3$};
\draw [red](4,1)--(6,1);
\draw [red](4,1)--(5,2.73);
\draw [red](6,1)--(5,2.73);
\node[font=\tiny, below] at (4,1) {$\alpha_1$};
\node[font=\tiny, below] at (6,1) {$\alpha_2$};
\node[font=\tiny, above] at (5,2.73) {$\alpha_3$};
\end{tikzpicture}
\end{figure}

Furthermore we assume $\tau_j:=\frac{Q(\alpha_j)}{P'(\alpha_j)}\in\mathbb{Q}$. The latter assumption guarantees the formal solution $\tilde{f}$ being slight, while the former assumption implies there will be three independent solutions.

Under the previous assumptions we prove that the Borel transformed solution $\hat{f}(\zeta_j)$ is hypergeometric, $\zeta_j=\zeta-\alpha_j$.

\begin{prop}
Let $P(\lambda)=\lambda^3+a_2\lambda^2+a_1\lambda+a_0$, $Q(\lambda)=b_2\lambda^2+b_1\lambda+b_0$ and $R(\tfrac{1}{z})=\tfrac{c_1}{z}+\tfrac{c_2}{z^2}$ satisfying the previous assumptions. Then 
\begin{itemize}
\item if $|\alpha_1-\alpha_2|<|\alpha_2-\alpha_3|<|\alpha_1-\alpha_3|$
\color{red}
\begin{align}
\hat{f}(\zeta_1)&=\left(1-\frac{\zeta_1}{\alpha_2-\alpha_1}\right)^{c-1} \, {}_2F_1\left(\mathbf{m};\mathbf{n};1-\frac{\zeta_1}{\alpha_2-\alpha_1} \right)\\
\hat{f}(\zeta_2)&=\left(1-\frac{\zeta_2}{\alpha_1-\alpha_2}\right)^{c-1} \, {}_2F_1\left(\mathbf{m};\mathbf{n};1-\frac{\zeta_2}{\alpha_1-\alpha_2} \right)\\
\hat{f}(\zeta_3)&=\left(1-\frac{\zeta_1}{\alpha_2-\alpha_1}\right)^{c-1} \, {}_2F_1\left(\mathbf{m};\mathbf{n};1-\frac{\zeta_3}{\alpha_2-\alpha_1} \right)
\end{align}
\color{black}
\item if $|\alpha_1-\alpha_2|<|\alpha_2-\alpha_3|=|\alpha_1-\alpha_3|$
\color{red}
\begin{align}
\hat{f}(\zeta_1)&=\left(1-\frac{\zeta_1}{\alpha_2-\alpha_1}\right)^{c-1} \, {}_2F_1\left(\mathbf{m};\mathbf{n};1-\frac{\zeta_1}{\alpha_2-\alpha_1} \right)\\
\hat{f}(\zeta_2)&=\left(1-\frac{\zeta_2}{\alpha_1-\alpha_2}\right)^{c-1} \, {}_2F_1\left(\mathbf{m};\mathbf{n};1-\frac{\zeta_2}{\alpha_1-\alpha_2} \right)\\
\hat{f}(\zeta_3)&=?
\end{align}
\color{black}
\item if $|\alpha_1-\alpha_2|=|\alpha_2-\alpha_3|=|\alpha_1-\alpha_3|$ 
\color{red}
\begin{align}
\hat{f}(\zeta_1)&=\left(1-\frac{\zeta_1}{\alpha_2-\alpha_1}\right)^{c-1} \, {}_3F_2\left(\mathbf{m};\mathbf{n};1-\frac{\zeta_1}{\alpha_2-\alpha_1} \right)\\
\hat{f}(\zeta_2)&=\left(1-\frac{\zeta_2}{\alpha_1-\alpha_2}\right)^{c-1} \, {}_3F_2\left(\mathbf{m};\mathbf{n};1-\frac{\zeta_2}{\alpha_1-\alpha_2} \right)\\
\hat{f}(\zeta_3)&=\left(1-\frac{\zeta_3}{\alpha_2-\alpha_1}\right)^{c-1} \, {}_3F_2\left(\mathbf{m};\mathbf{n};1-\frac{\zeta_3}{\alpha_2-\alpha_1} \right)
\end{align}

\end{itemize}

\color{black}
where the coefficients $\mathbf{m}=(m_1,m_2,m_3),\mathbf{n}=(n_1,n_2)$ depend on the parameter of $P,Q,R$. 
\end{prop}

\begin{proof}
The strategy is analogous to the one for second order ODEs. Hence, we first consider the Borel transform of equation \ref{eq3}: 
\begin{multline}\label{borel 3}
P(-\zeta)\hat{f}+\int_0^\zeta Q(-\zeta')\hat{f}(\zeta')d\zeta'+c_1\int_0^\zeta(\zeta-\zeta')\hat{f}(\zeta')d\zeta'+\frac{c_2}{2}\int_0^\zeta(\zeta-\zeta')^2\hat{f}(\zeta')d\zeta'=0
\end{multline}

Although equation \eqref{borel 3} is an integral equation, since $\tilde{f}$ is slight, $\hat{f}$ is a solution of \eqref{borel 3} if and only if it is a solution of the third order ODE obtained by differentiating \eqref{borel 3} three times: 
\begin{multline}\label{borel_ODE 3}
\big[P(-\zeta)\partial_\zeta^3+(-3P'(-\zeta)+Q(-\zeta))\partial_\zeta^2+(3P''(-\zeta)-2Q'(-\zeta)+c_1)\partial_\zeta+(-P'''(-\zeta)+Q''(-\zeta)+c_2)\big]\hat{f}=0
\end{multline}

Notice that

\begin{align*}
P(-\zeta)&=(\zeta-\alpha_1)(\zeta-\alpha_2)(\zeta-\alpha_3)\\
-P'(-\zeta)&=(\zeta-\alpha_2)(\zeta-\alpha_3)+(\zeta-\alpha_1)(\zeta-\alpha_3)+(\zeta-\alpha_2)(\zeta-\alpha_1)\\
P''(-\zeta)&=6\zeta-2\alpha_1-2\alpha_2-2\alpha_3
\end{align*}

We now study separately the three scenarios:
\begin{itemize}
\item if $\ell=|\alpha_1-\alpha_2|<|\alpha_2-\alpha_3|<|\alpha_1-\alpha_3|$ we claim that \eqref{borel_ODE 3} is ??
\item if $\alpha_2=\alpha_3+\ell,\, \alpha_1=\alpha_3-\ell$ we claim that \eqref{borel_ODE 3} is a ??
\color{Aquamarine}
Change coordinates $y=\frac{\zeta-\alpha_1}{\ell}$:
\begin{align*}
P(-\zeta)\partial_\zeta^3&\longrightarrow -\ell^3y(\ell y+\alpha_1-\alpha_2)(\ell y+\alpha_1-\alpha_3)\frac{1}{\ell^3}\partial_y^3=y(1-y^2)\partial_y^3\\
(-3P'(-\zeta)+Q(-\zeta))\partial_\zeta^2&\longrightarrow (3\ell^{2}(3y^2-1)+Q(-\ell y-\alpha_1))\frac{1}{\ell^2}\partial_y^2\\
(3P''(-\zeta)-2Q'(-\zeta)+c_1)\partial_\zeta&\longrightarrow(-18\ell y-2Q'(-\ell y-\alpha_1)+c_1)\frac{1}{\ell}\partial_y\\
-P'''(-\zeta)+Q''(-\zeta)+c_2&\longrightarrow -6+2b_2+c_2
\end{align*}

Change coordinates $s=y^2$:
\begin{align*}
y(1-y^2)\partial_y^3&\longrightarrow 8(1-s)s^2\partial_s^3+12(1-s)s\partial_s^2\\
(3\ell^{2}(3y^2-1)+Q(-\ell y-\alpha_1))\frac{1}{\ell^2}\partial_y^2&\longrightarrow \big[6(3s-1)+2(b_2s+\frac{2b_2\alpha_1-b_1}{\ell}y+\frac{b_2\alpha_1^2-\alpha_1b_1+b_0}{\ell^2})\big](\partial_s+2s\partial_s^2)\\
(-18\ell y-2Q'(-\ell y-\alpha_1)+c_1)\frac{1}{\ell}\partial_y&\longrightarrow 4(2b_2-9)s\partial_s+2\frac{4b_2\alpha_1-2b_1+c_1}{\ell}y\partial_s
\end{align*}

\begin{multline}
\Big[8(1-s)s^2\partial_s^3+\big[4(6+b_2)s+4\frac{2b_2\alpha_1-b_1}{\ell}y+4\frac{b_2\alpha_1^2-\alpha_1b_1+b_0}{\ell^2}\big]s\partial_s^2\Big.\\
\Big.\big[-2(9-2b_2)s-2(3-2b_2)+2\frac{b_2\alpha_1^2-\alpha_1b_1+b_0}{\ell^2}+2\frac{6b_2\alpha_1-3b_1+c_1}{\ell}y\big]\partial_s-6+2b_2+c_2\Big]\hat{f}=0
\end{multline}

\begin{align*}
\frac{2b_2\alpha_1-b_1}{\ell}y+\frac{b_2\alpha_1^2-\alpha_1b_1+b_0}{\ell^2}=2n_1+2n_2-2\\
6+b_2=-2(3+m_1+m_2+m_3)\\
\big[-(3-2b_2)+2(\frac{b_2\alpha_1^2-\alpha_1b_1+b_0}{\ell^2})+\frac{8b_2\alpha_1-4b_1+c_1}{\ell}y\big]=4n_1n_2\\
(9-2b_2)=4(1+m_1+m_2+m_3+m_1m_2+m_1m_3+m_2m_3)\\
-6+2b_2+c_2=8m_1m_2m_3
\end{align*}
\color{black}
\item if $|\alpha_1-\alpha_2|=|\alpha_2-\alpha_3|=|\alpha_1-\alpha_3|=\ell$ we claim that \eqref{borel_ODE 3} is a generalized hypergeometric equation. Without loss of generality we assume $\alpha_2=\alpha_1+\ell$ and $\alpha_3=\alpha_1+\frac{\ell}{2}+i\frac{\sqrt{3}}{2}\ell$.

Change coordinates $y=\frac{\zeta-\alpha_1}{\ell}+\frac{\sqrt{3}+i}{2}$, then 
\begin{align*}
P(-\zeta)\partial_\zeta^3&\longrightarrow -\ell y(\ell y+\alpha_1-\alpha_2)(\ell y+\alpha_1-\alpha_3)\frac{1}{\ell^3}\partial_y^3=y(1-y^2)\partial_y^3\\
(-3P'(-\zeta)+Q(-\zeta))\partial_\zeta^2&\longrightarrow (3\ell^{2}(3y^2-1)+Q(-\ell y-\alpha_1))\frac{1}{\ell^2}\partial_y^2\\
(3P''(-\zeta)-2Q'(-\zeta)+c_1)\partial_\zeta&\longrightarrow(-18\ell y-2Q'(-\ell y-\alpha_1)+c_1)\frac{1}{\ell}\partial_y\\
-P'''(-\zeta)+Q''(-\zeta)+c_2&\longrightarrow -6+2b_2+c_2
\end{align*} 
\end{itemize}

\end{proof}

\bibliographystyle{plain}
\bibliography{airy-resurgence}
\end{document}