### 2024-01-17 to-do.tex
Sync before meeting

\documentclass{article}

\title{To-do}
\date{}

\begin{document}
\maketitle
\section{Intro}
\begin{itemize}
\item Finish new version of ``Why does Borel summation work for solutions of level 1 ODEs?''
\end{itemize}
\section{The Laplace and Borel transforms}
\begin{itemize}
\item Finish ``Regularity and decay properties''
\end{itemize}
\section{Proof of main results}
\begin{itemize}
    \item Finish new version of ''Borel regularity for ODEs''
\end{itemize}
\section{Examples}
\begin{itemize}
\item Revise Borel regularity argument for Airy-Lucas
\item Rewrite thimble projection reasoning for Airy-Lucas functions
\item Explain equation of tringular cantilver
\end{itemize}
\section{Figures}
For some reason all figures move at the end of the document. Moreover I'd prefer to add the picture of the Borel plane as translation surface as tikzpicture and not as pdf.
\end{document}