### 2024-05-03 sima/correspondence/sima-v2-changes.tex
Drop complex tau scratch work

\documentclass{letter}

\usepackage{url}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{tikz}

% include signature image. based on suggestion from TeX.SE user Werner
%   https://tex.stackexchange.com/a/148458/6934
%%\makeatletter
%%\newcommand*{\signatureimg}[1]{\def\fromimg{#1}}
%%\renewcommand{\closing}[1]{\par\nobreak\vspace{\parskip}%
%%  \stopbreaks
%%  \noindent
%%  \ifx\@empty\fromaddress\else
%%  \hspace*{\longindentation}\fi
%%  \parbox{\indentedwidth}{\raggedright
%%       \ignorespaces #1\\%
%%       \includegraphics[height=6\medskipamount]{\fromimg}\\%
%%       \ifx\@empty\fromsig
%%           \fromname
%%       \else \fromsig \fi\strut}%
%%   \par}
%%\makeatother

\address{Veronica Fantini \\
Institut des Hautes \'{E}tudes Scientifiques \\
Le Bois-Marie, 35 route de Chartres \\
F-91440 Bures-sur-Yvette, France}

\signature{\vspace{-12mm}Veronica Fantini \\
\url{fantini@ihes.fr} \\[6pt]
Aaron Fenyes \\
\url{aaron.fenyes@fareycircles.ooo}}

%%\signatureimg{placeholder-sig.pdf}

\date{March \textcolor{violet}{--}, 2023}

\begin{document}
\begin{letter}{}
\begin{tikzpicture}[remember picture,overlay]
\node[xshift=4.2cm, yshift=-3.5cm] at (current page.north west) {\pgftext{\includegraphics[height=4.02cm]{ihes-logo}}};
\end{tikzpicture}
\opening{Dear Prof. Lipton:}

\color{violet}
Reviewer~2:
\begin{itemize}
    \item We corrected the sign error that reviewer~2 noticed on line~275.
    \item We corrected the factorial $j!$ as reviewer~2 suggested on line~706.
     \item We corrected the factorial $j!$ as reviewer~2 suggested on line~730. We rewrote the decomposition of $Q$. 
     \item We corrected the sign error that reviewer~2 noticed on line~767.
      \item We corrected the typo that reviewer~2 noticed on line~781.
\end{itemize}
Reviewer~1:
\begin{itemize}
\item Reviewer~1 asked us to write more concisely and more formally. Where these goals conflict, we have prioritized conciseness. For example, we have continued to use ``small enough'' instead of ``sufficiently small.''
\item There is no universal convention for naming the coordinates on the spaces involved in the Laplace transform. Conventions we have seen include:
\begin{center}
\begin{tabular}{c|c|l}
Frequency varaible & Position variable & Source \\ \hline
$z$ & $\zeta$ & Mitschi \& Sauzin \\
$s$ & $t$ & Schiff \\
$\hbar$ & $s$ & Mari\~{n}o \\
\end{tabular}
\end{center}
From a physical perspective, the Laplace transform turns functions of position or time into functions of spatial or temporal frequency. The terms ``frequency'' and ``position'' thus unambiguously distinguish the two spaces involved in the Laplace transform. We use these terms to remind the reader which variable-naming convention we are using.
\item We have removed all contractions.
\item We have removed the phrase ``tells us.''
\item We use the phrase ``over all'' to emphasize that we are describing a uniform bound. This is now explained explicitly in the section ``Notation for uniform bounds.''
\item We have changed our notation for uniform bounds to require the absolute value.
\item the seminorm is a seminorm, because we can choose $K=\{\mathsf{pt}\}$
\end{itemize}
\color{orange}
For $\tau\in\mathbb{C}$
\begin{equation}
    \left(\frac{\zeta(a)}{\zeta(b)}\right)^\tau=\exp\left(\Re(\tau) \textcolor{magenta}{\log} \frac{\zeta(a)}{\zeta(b)}+i\Im(\tau)\textcolor{magenta}{\log} \frac{\zeta(a)}{\zeta(b)}\right)
\end{equation}
so if $\Re(\tau)>0$ and it is bigger than $\Im(\tau)$ the first term dominates in the limit $\zeta\to0$...but maybe we don't need ``bigger than $\Im(\tau)$", as ``we can rotate". I'm not very confortable with rotations, so probably I'll not add this complication. I'm also afraid we should be more carefull with these mooving domains, if we allow rotations. So maybe better not make things too complicated for the sake of generality. We have a good list of examples that fit in our pictures. 
\color{black}
\color{black}
\closing{Best regards,}
\end{letter}
\end{document}